---
title: "Catch model"
author: "Collin Edwards"
editor: visual
---

## Todo

- run with angler minutes
- problems with 2021 data, run without that
- incorporate information on angler type
- Model jacks!!

## Purpose

Here we develop a model to predict fish encounters in the creel interviews, and identify when and how observations fail to match the expectations of our fitted model.

In v2, I am revisiting this work starting June 2025.

## Setup

### libraries

```{r}
library(here)
library(tidyverse)
library(lme4)
library(mgcv)
library(patchwork)
library(gt)
library(gtExtras)
library(gratia)
library(DHARMa)
```

### functions

```{r}
doy_2md=function(i){
  ymd=as.Date(i-1, origin="2019-01-01")
  return(format(ymd, "%b %d"))
}

facet_names <- c(`FALSE` = "Reported no kept fish",
                 `TRUE` = "Reported 1+ kept fish")

```

### Reading in the data

```{r}
## rerun the following to recreate the clean data
if(FALSE){
  source(here("scripts/prepare_creel_data.R"))
}

raw = read_csv(here("cleaned_data/key_dataframes/creel_interview_catch_withzeros.csv")) 
## Note: problem with col 4 being interpreted as a T/F instead of char

dat = raw |> 
  filter(species == "Chinook")

dat.sum = dat |> 
  filter(fin_mark %in% c("UM", "AD")) |> 
  group_by(interview_id, event_date, year, month, fishing_duration_minutes,
           angler_count, angler_minutes, trip_status, angler_type, trip_guided, boat_type, boat_used, fishing_start_time, fishing_end_time) |> 
  summarize(um_adult = sum(fish_count[fin_mark == "UM" & life_stage == "Adult"]),
            ad_adult = sum(fish_count[fin_mark == "AD" & life_stage == "Adult"]),
            released_adult = sum(fish_count[fate == "Released" & life_stage == "Adult"]),
            kept_adult = sum(fish_count[fate == "Kept" & life_stage == "Adult"]),
            um_jack = sum(fish_count[fin_mark == "UM" & life_stage == "Jack"]),
            ad_jack = sum(fish_count[fin_mark == "AD" & life_stage == "Jack"]),
            released_jack = sum(fish_count[fate == "Released" & life_stage == "Jack"]),
            kept_jack = sum(fish_count[fate == "Kept" & life_stage == "Jack"])
  ) |> 
  ungroup() |> 
  mutate(um_total = um_jack + um_adult,
         ad_total = ad_jack + ad_adult,
         released_total = released_adult + released_jack,
         kept_total = kept_adult + kept_jack,
         had.ad = ad_total > 0,
         had.um = um_total > 0,
         had.release = released_total > 0,
         had.kept = kept_total > 0)

## for now, ignore jacks

dat.fit = dat |> 
  # filter(life_stage == "Adult") |> 
  filter(fin_mark %in% c("UM", "AD")) |> 
  group_by(interview_id, life_stage, fin_mark, event_date, year, month, fishing_duration_minutes, 
           angler_count, angler_minutes, trip_status, angler_type, trip_guided, boat_type, boat_used,
           fishing_start_time, fishing_end_time) |>   
  summarize(fish_count = sum(fish_count)) |> 
  ungroup() |> 
  ## need to collapse the data down a bit
  left_join(dat.sum |> 
              select(interview_id, had.ad, had.um, had.release, had.kept, um_total, ad_total),
            by = "interview_id") |> 
  mutate(doy = yday(event_date),
         fin_mark = as.factor(fin_mark),
         had.ad = as.factor(had.ad),
         had.um = as.factor(had.um),
         had.release = as.factor(had.release),
         had.kept = as.factor(had.kept),
         yearfac = as.factor(year)) |> 
  filter(!is.na(fishing_duration_minutes)) |> 
  mutate(round_cat1 = case_when(
    fish_count >= 3 & fish_count <= 7 ~ "3-7",
    fish_count >= 8 & fish_count <= 12 ~ "8-12",
    fish_count >= 13 ~ "13+",
  )) |> 
  mutate(round_cat1 = factor(round_cat1, levels = c("3-7", "8-12", "13+"))) |> 
  mutate(round_cat2 = case_when(
    fish_count >= 3 & fish_count <= 7 ~ "3-7",
    fish_count >= 8 ~ "8+"
  )) |> 
  mutate(round_cat2 = factor(round_cat2, levels = c("3-7", "8+")))
```

Note that we're skipping the \~500 (out of \~65,000) observations with no fishing time. These are entirely from incomplete trips in 2021, which can't be compared apples-to-apples with completed trips anyways.

#### Removing 2021

The 2021 data we have available doesn't match the underlying datasheets. Until BDS fixes this, we will use only 2022 and 2023 data.

```{r}
dat.sum = dat.sum |> 
  filter(year != 2021)

dat.fit = dat.fit |> 
  filter(year != 2021)
```


::::: panel-tabset

## Looking at the data

The number of fish caught should vary based on time of year and time spent fishing, but we can still start with some basic summaries of the number of reported unmarked fish.

```{r}
dat.sum |> 
  count(um_total) |> 
  rename(unmarked_encounters = um_total,
         interviews = n) |> 
  gt()
```

We can look at the breakdown of angler minutes per trip to see if the highest UM encounter trips

```{r}
ggplot(dat.sum, aes(x = angler_minutes, y = um_total, col = um_total==10))+
  geom_jitter(width = 0, height = 0.3)+theme_bw()+
  labs(x = "angler minutes (duration x # of anglers)", y = "Unmarked encounters",
       title = "Unmarked encounters by fishing time")
```

It can be hard to see what the average minutes per catch look like, so here they are in table form

```{r}
dat.sum |> 
  group_by(um_total) |> 
  summarize(n = n(),
            mean_angler_minutes = mean(angler_minutes, na.rm = T)) |> 
  rename(unmarked_encounters = um_total,
         interviews = n) |> 
  gt()
```

It's also relevant to look at what time of year this occured in.

```{r}
dat.sum |> 
  mutate(doy = yday(event_date)) |> 
  ggplot(aes(x = doy, y = um_total, col = um_total==10))+
  geom_jitter(width = 0, height = 0.3)+theme_bw()+
  labs(y = "Unmarked encounters",
       title = "Unmarked encounters by season")+
  scale_x_continuous(labels = doy_2md)
```

They certainly ARE the time of year when unmarked encounters were highest.

### Minutes per catch

We can look at how realistic some of this data is by calculating the average angler minutes per catch, and see if some of the observations are not realistic.
Trips with no catch will be treated as NAs for this calculation.

```{r}
dat.sum = dat.sum |> 
  mutate(catch_per_hour = (um_total + ad_total)/angler_minutes * 2,
         minutes_per_catch = angler_minutes / (um_total + ad_total),
         minutes_per_catch = if_else(is.infinite(minutes_per_catch),
                                     NA,
                                     minutes_per_catch))

hist(dat.sum$minutes_per_catch, breaks = 500)

hist(dat.sum$minutes_per_catch, breaks = 100, xlim = c(0, 500))
```

Let's look at the top 20 fastest catching surveys:

```{r}
dat.sum |> 
  arrange((minutes_per_catch)) |> 
  select(minutes_per_catch, um_total, ad_total, fishing_duration_minutes , angler_count, boat_used, event_date, fishing_start_time, fishing_end_time) |> 
  head(20) |> 
  gt() |> 
  gt_highlight_rows(rows = um_total == 5)

```

We have an impossible data point with identical start and end time. Other than that, we are looking at generally seems feasible. Further, the data is not dominated by surveys that are primarily high UM encounters. That said, our two highest UM reported (5 UM encounters) are in the top 20 most efficient catch trips, with averages of 30 and 40 angler-minutes per fish for those trips. Trips with high UM catch must necessarily have high catch rates (or huge surpluses of angler-minutes), but this might border on the unreasonable?



```{r}
dat.plot = dat.sum |> 
  mutate(doy = yday(event_date))

dat.plot |> 
  ggplot(aes(x = doy, y = catch_per_hour))+
  geom_jitter(width = 0.2, height = 0)+
  geom_text(data = dat.plot |> filter((catch_per_hour) > 1.3),
            aes(y = catch_per_hour, x = doy, label = um_total),
            col = 'blue', nudge_x = 1, nudge_y = -.08)+
  facet_wrap(.~ year, ncol = 1)+ 
  scale_x_continuous(labels = doy_2md)+
  labs(x = "", y = "Chinook caught per hour")+
  labs(title = "Chinook catch per hour",
       subtitle = "black numbers = total unmarked catch")+
  theme_bw()


dat.plot |> 
  ggplot(aes(x = doy, y = catch_per_hour, col = as.factor(year)))+
  geom_jitter(width = 0.2, height = 0)+
  labs(x = "", y = "Chinook caught per hour")+
  scale_x_continuous(labels = doy_2md)+
  labs(title = "Chinook catch per hour",
       subtitle = "date jittered +/- 5 hours very slightly for readability")+
  theme_bw()

dat.plot |> 
  ggplot(aes(x = doy, y = catch_per_hour, col = as.factor(year)))+
  geom_smooth()+
  labs(x = "", y = "Chinook caught per hour")+
  scale_x_continuous(labels = doy_2md)+
  labs(title = "Chinook catch per hour",
       subtitle = "date jittered +/- 5 hours very slightly for readability")+
  theme_bw()


```

Request for "density" plot of above, shaded by year. Density plot is analogous to histogram, so going to try simple smooth approach


```{r}
temp = dat.sum |> select(boat_used, minutes_per_catch) |> 
  mutate(minutes_per_catch = coalesce(minutes_per_catch, Inf))

quantile(temp$minutes_per_catch[temp$boat_used=="No"], c(0.001, 0.01, 0.025, 0.05, 0.1, 0.2))
quantile(temp$minutes_per_catch[temp$boat_used=="Yes"], c(0.001, 0.01, 0.025, 0.05, 0.1, 0.2))
temp |> 
  group_by(boat_used) |> 
  summarize(prop_at_least_hourly = mean(minutes_per_catch <= 60))
temp |> 
  group_by(boat_used) |> 
  summarize(prop_at_least_once_per_90 = mean(minutes_per_catch <= 90))

```

### hypothetical filtering by efficiency

```{r}
efficiency = seq(1:120)
res.ls = list()
for(i in 1:length(efficiency)){
  cur.counts = dat.sum |> 
    filter(minutes_per_catch > efficiency[i]) |> 
    count(boat_used)
  cur.loss = dat.sum |> 
    filter(minutes_per_catch <= efficiency[i]) |> 
    count(boat_used)
  res.ls[[i]] = tibble(threshold_efficiency_minutes = efficiency[i],
                       boat_surveys_remaining = cur.counts |> filter(boat_used == "Yes") |> pull(n),
                       bank_surveys_remaining = cur.counts |> filter(boat_used == "No") |> pull(n),
                       boat_surveys_lost = cur.loss |> filter(boat_used == "Yes") |> pull(n),
                       bank_surveys_lost = cur.loss |> filter(boat_used == "No") |> pull(n),
                       )
}
res.df = do.call(bind_rows, res.ls)


ggplot(res.df, aes(x = threshold_efficiency_minutes, y = boat_surveys_remaining))+
  geom_path()+
  ylim(c(0, NA))+
  ggtitle("Boat surveys with thresholding")

ggplot(res.df, aes(x = threshold_efficiency_minutes, y = bank_surveys_remaining))+
  geom_path()+
  ylim(c(0, NA))+
  ggtitle("Bank surveys with thresholding")

ggplot(res.df, aes(x = threshold_efficiency_minutes, y = bank_surveys_lost + boat_surveys_lost))+
  geom_path()+
  ggtitle("Total surveys lost with thresholding")


ggplot(res.df, aes(x = threshold_efficiency_minutes, y = (bank_surveys_lost + boat_surveys_lost)/nrow(dat.sum)))+
  geom_path()+
  ggtitle("Surveys lost with threshold-based filtering")+
  scale_y_continuous(labels = scales::percent)+
  labs(y = "% of surveys removed",
       x = "Threshold minutes per catch")+
  theme_bw(base_size = 14)



```



### Catch by duration


```{r}
dat.sum |> 
  mutate(total_encounters = um_total+ad_total) |> 
  ggplot(aes(x = angler_minutes, y = total_encounters, col = boat_used))+
  geom_point()+
  geom_smooth()+
  facet_wrap(.~ boat_used)+
  labs(title = "Encounters by time angling")

```

### Catch by start time

```{r}
dat.sum |>
  mutate(total_encounters = um_total+ad_total) |> 
  ggplot(aes(x = fishing_start_time, y = catch_per_hour, col = boat_used))+
  geom_point()+
  geom_smooth()+
  facet_wrap(.~ boat_used)+
  labs(y = "catch per hour",
       title = "catch rate as a function of start time")+
  theme_bw()
```

### AD catch by boat_used

```{r}
dat.fit |> filter(life_stage == "Adult", fin_mark == "AD") |> 
  count(ad_total, boat_used) |> pivot_wider(names_from = boat_used, values_from = n) |> 
  gt() |> 
  tab_caption("# of surveys by marked encounters, split by boat_used")
  
```




## Simple gam model

We fit a generalized additive model with fixed effect of fin_mark and fishing duration, random effect of year, and smooth across day of year that differs by fin_mark. We assume a negative binomial distribution (more on that later).


::: panel-tabset
```{r}
## GAM model -------------------------------------------

dat.fit = dat.fit |> 
  mutate(mark_stage = as.factor(paste0(fin_mark, " X ", life_stage)))

out = gam(fish_count ~ mark_stage +
            s(doy, by = mark_stage, k = 20) + 
            s(yearfac, bs = 're') + 
            angler_minutes,
          method = "REML",
          family = "nb",
          data = dat.fit)

# draw(out)

gp.seasonality = conditional_values(
  out,
  condition = c("doy", "mark_stage", "yearfac")
) |> 
  draw()+
  scale_x_continuous(limits = c(200, 300),
                     labels = doy_2md)+
  scale_y_continuous(limits = c(0, .5))+
  facet_wrap(.~ yearfac)+
  labs(x = "",
       y = "Predicted encounters per hour",
       title = "Seasonality of encounters, same shape per year")+
  theme_bw(base_size = 14)

gp.seasonality

```

```{r}
## GAM model -------------------------------------------

dat.fit = dat.fit |> 
  mutate(mark_stage = as.factor(paste0(fin_mark, " X ", life_stage)))

out.yearly = gam(fish_count ~ 
                   te(doy, mark_stage, yearfac, bs = c('tp', 'fs', 'fs')) + 
                   # s(yearfac, bs = 're') + 
                   angler_minutes,
                 method = "REML",
                 family = "nb",
                 data = dat.fit)

# draw(out)

gp.seasonality = conditional_values(
  out.yearly,
  condition = c("doy", "mark_stage", "yearfac")
) |> 
  draw()+
  scale_x_continuous(limits = c(200, 300),
                     labels = doy_2md)+
  scale_y_continuous(limits = c(0, .5))+
  facet_wrap(.~ yearfac)+
  labs(x = "",
       y = "Predicted encounters per hour",
       title = "Seasonality of encounters, different shapes each year")+
  theme_bw(base_size = 14)

gp.seasonality

## Checking results, as they don't match Evan's bar plots

dat.fit$pred = as.numeric(predict(out.yearly, newdata = dat.fit, type = "response"))

dat.fit |> 
  mutate(doy = yday(event_date)) |> 
ggplot(aes(x = doy, y = pred, col = mark_stage))+
  geom_jitter(width = 0, height = 0.2)+
  facet_wrap(.~ year)+
  scale_x_continuous(labels = doy_2md)

## Okay, what about a "generic" trip

dat.pred = expand_grid(doy = seq(min(dat.fit$doy), max(dat.fit$doy), by = 1),
                       mark_stage = factor("AD X Adult", levels = levels(dat.fit$mark_stage)),
                       yearfac = unique(dat.fit$yearfac),
                       angler_minutes = 120
                       )

dat.pred$count_predicted = as.numeric(predict(out.yearly, newdata = dat.pred, type = "response"))

dat.pred |> 
ggplot(aes(x = doy, y = count_predicted, col = mark_stage))+
  geom_point()+
  facet_wrap(.~ yearfac)+
  scale_x_continuous(labels = doy_2md)+
  theme_bw(base_size = 13)+
  labs(x = "", title = "Daily predictions for 2 angler-hours")

dat.pred |> 
ggplot(aes(x = doy, y = count_predicted, col = yearfac, fill = yearfac))+
  geom_polygon(alpha = 0.2)+
  geom_path()+
  scale_x_continuous(labels = doy_2md)+
  theme_bw(base_size = 13)+
  labs(x = "", title = "Daily predictions for 2 angler-hours")


dat.fit |> 
  group_by(year, doy) |> 
  summarise(angler_hours = sum(angler_minutes)/60) |> 
ggplot(aes(x = doy, y = angler_hours, col = as.factor(year), fill = as.factor(year)))+
  # geom_col(alpha = 0.2)+
  geom_smooth()+
  scale_x_continuous(labels = doy_2md)+
  theme_bw(base_size = 13)

```


```{r}
dat.fit |> 
  group_by(month, year) |> 
  summarize(ad_per_hour = sum(ad_total) / sum(angler_minutes) * 60) |> 
  ungroup() |> 
  ggplot(aes(x = month, y = ad_per_hour, fill = as_factor(year)))+
  geom_col()+
  facet_wrap(.~ year)+
  theme_bw(base_size = 13)

gp.ad = dat.fit |> 
  group_by(month, year) |> 
  summarize(ad = sum(ad_total)) |> 
  ungroup() |> 
  ggplot(aes(x = month, y = ad, fill = as_factor(year)))+
  geom_col()+
  facet_wrap(.~ year)+
  labs(title = "Fish encountered", 
       y = "total ad encounters")+
  theme_bw(base_size = 13)

gp.time = dat.fit |> 
  group_by(month, year) |> 
  summarize(angler_minutes = sum(angler_minutes)) |> 
  ungroup() |> 
  ggplot(aes(x = month, y = angler_minutes, fill = as_factor(year)))+
  geom_col()+
  facet_wrap(.~ year)+
  labs(y = "total angler minutes",
       title = "Time spent")+
  theme_bw(base_size = 13)

gp.ad / gp.time
```


What is our total time breakdown by hour?

```{r}
dat.fit |> 
  group_by(year) |> 
  summarize(angler_hours = sum(angler_minutes)/60,
            ad_total = sum(ad_total)) |> 
  gt()
```

```{r}
dat.fit |> 
  group_by(year) |> 
  summarize(angler_hours = sum(angler_minutes)/60) |> 
  gt()
```

Angler minutes across seasons
```{r}
ggplot(dat.fit, aes(x = doy, y = angler_minutes))+
  geom_jitter(width = 0.2, height = 0.2)+
  scale_x_continuous(labels = doy_2md)+
  facet_wrap(.~ yearfac)
```


```{r}
out.yearly.boatly = gam(fish_count ~ 
                          te(doy, mark_stage, yearfac, bs = c('tp', 'fs', 'fs')) + 
                          boat_used +
                          angler_minutes,
                        method = "REML",
                        family = "nb",
                        data = dat.fit)

out.yearly.boatly2 = gam(fish_count ~ 
                           te(doy, mark_stage, yearfac, bs = c('tp', 'fs', 'fs')) + 
                           angler_minutes:boat_used,
                         method = "REML",
                         family = "nb",
                         data = dat.fit)

out.yearly.boatly3 = gam(fish_count ~ 
                           te(doy, mark_stage, yearfac, bs = c('tp', 'fs', 'fs')) + 
                           angler_minutes*boat_used,
                         method = "REML",
                         family = "nb",
                         data = dat.fit)
```

```{r}
AIC(out, out.yearly, out.yearly.boatly, out.yearly.boatly2, out.yearly.boatly3)
```

We see the most support for a model that allows smooths to vary by year and accounts for boat effect -- both as a fixed effect (boat anglers caught more regardless of trip length) and interacting with angler minutes (After accounting for the general boat effect, boat anglers gained less by having longer trips than non-boat anglers).

Out of curiosity, how much does boat effect matter?

```{r}
coef(out.yearly.boatly)[2:3]
```

That's a huge difference. How many minutes of fishing does boating count as?

```{r}
temp = coef(out.yearly.boatly)[2]/coef(out.yearly.boatly)[3]
```

A lot! `r round(temp/60, 1)` hours! That means the model predicts that a single angler on a boat fishing for an hour would have the same catch as a group of 5 anglers fishing for 6 hours.

Let's look at the partial sum of predictors for a boat-user who fished for 60 minutes:
```{r}
sum(coef(out.yearly.boatly)[2:3]*c(1,60))
```


Quick look at benefits of boating and time on the water:

```{r}
dat.pred.1 = expand_grid(doy = mean(dat.fit$doy),
                         yearfac = unique(dat.fit$yearfac)[1],
                         mark_stage = dat.pred$mark_stage[1],
                         angler_minutes = seq(min(dat.fit$angler_minutes), max(dat.fit$angler_minutes),
                                              by = 1),
                         boat_used = unique(dat.fit$boat_used))
dat.pred.1$predicted_count = as.numeric(predict(out.yearly.boatly3, newdata = dat.pred.1, type = "response"))

ggplot(dat.pred.1, aes(x = angler_minutes, y = predicted_count))+
  geom_path()+
  geom_rug(data = dat.fit |> filter(mark_stage == "AD X Adult"), aes(x = angler_minutes, y = fish_count))+
  facet_wrap(.~ boat_used)+
  labs(y = "predicted AD encounters",
       title = "angler efficiency by time and boat status")
```



### model diagnostics

#### Shared shape model

Examining the model that assumes the same smooth across doy in each year (seasonality is the same each year).

First we check the model diagnostics

```{r}
gam.check(out, rep = 500)
```

These don't look great, and may reflect a miss-specified family.

We use the Dharma package to look for model fit. Note that because this is simulation-based and we have MANY observations, the significance of tests for deviations are going to be inflated. I'm looking into solutions to this.

```{r}
simulationOutput = simulateResiduals(fittedModel = out, plot = F)
plot(simulationOutput)
testOutliers(simulationOutput, type = "bootstrap")
testDispersion(simulationOutput)
testZeroInflation(simulationOutput)
```


These seem fine -- we are not seeing any signs of problems

#### Separate shape model

Examining the model that assumes different smooths across doy in each year (seasonality can differ between years).

First we check the model diagnostics

```{r}
gam.check(out.yearly, rep = 500)
```

These are not informative, although the k' being twice the edf suggests we likely aren't running too little flexibility (although this can be hard to parse given the high dimensionality).

```{r}
simulationOutput = simulateResiduals(fittedModel = out.yearly, plot = F)
plot(simulationOutput)
testOutliers(simulationOutput, type = "bootstrap")
testDispersion(simulationOutput)
testZeroInflation(simulationOutput)
```

These seem fine -- we are not seeing any signs of problems

### calculating predicted distributions

```{r}
dat.fit$fish_count_predicted = as.numeric(predict(out, newdata = dat.fit, type = "response"))
dat.fit = dat.fit |>
  mutate(fish_count_predicted_int = round(fish_count_predicted))


## Looking at predicted vs observed distributions of gam model ==================

## Returning to our simple gam model: 

mus = predict(out, newdata = dat.fit, type = "response")
cur.theta = out$family$getTheta(TRUE)

# dnbinom(0:10, size = cur.theta, mu = mus[1])

foo <- function(x){dnbinom(0:30, size = cur.theta, mu = x)}
df = data.frame(mu = mus)|> 
  mutate(mat = do.call(rbind, purrr::map(mu, foo)))

df.pred = as.data.frame(df$mat)
#gros names for a sec:
names(df.pred) = 0:30  
## now combine with the data and use pivoting to put into handy form
df.pred <- dat.fit |> 
  select(-fish_count_predicted, -fish_count_predicted_int) |> 
  cbind(df.pred) |> 
  pivot_longer(cols = "0":"30",
               names_to = "predicted_count",
               values_to = "predicted_probability") |> 
  mutate(predicted_count = as.numeric(predicted_count))

## With this dataframe, we can now sum up the number of counts we would expect to see
## for marked and unmarked fish, *accounting* for the variability in negative binomials

freq.pred = df.pred |> 
  group_by(fin_mark, predicted_count) |> 
  summarize(expected_frequency = sum(predicted_probability)) |> 
  ungroup() |> 
  ## adding rounding categorizations
  mutate(round_cat1 = case_when(
    predicted_count >= 3 & predicted_count <= 7 ~ "3-7",
    predicted_count >= 8 & predicted_count <= 12 ~ "8-12",
    predicted_count >= 13 ~ "13+",
  )) |> 
  mutate(round_cat1 = factor(round_cat1, levels = c("3-7", "8-12", "13+"))) |> 
  mutate(round_cat2 = case_when(
    predicted_count >= 3 & predicted_count <= 7 ~ "3-7",
    predicted_count >= 8 ~ "8+"
  )) |> 
  mutate(round_cat2 = factor(round_cat2, levels = c("3-7", "8+")))

## basic plotting ================================

gp.fit = dat.fit |> 
  filter(fin_mark == "AD") |> 
  count(fish_count) |> 
  ggplot(aes(x = fish_count, y = n))+
  geom_col()+
  xlim(c(-1,11))+
  ggtitle("Observed counts of AD fish")+
  theme_bw(base_size = 14)

gp.pred = freq.pred |> 
  filter(fin_mark == "AD") |> 
  ggplot(aes(x = predicted_count, y = expected_frequency))+
  geom_col()+
  ggtitle("Predicted counts of AD fish")+
  xlim(c(-1,11))+
  theme_bw(base_size = 14)

freq.pred |> 
  ggplot(aes(x = predicted_count, y = expected_frequency))+
  geom_col()+
  ggtitle("")+
  xlim(c(-1,11))+
  facet_wrap(. ~fin_mark)+
  labs(y = "Expected # of interviews",
       x = "reported encounters")+
  theme_bw(base_size = 18)


## AD marked

pred.partial = freq.pred |> 
  filter(fin_mark == "AD") |> 
  group_by(predicted_count) |> 
  summarize(expected_frequency = sum(expected_frequency)) |> 
  ungroup() |> 
  mutate(scenario = "prediction") |> 
  rename(fish_count = predicted_count, frequency = expected_frequency)

obs.partial = dat.fit |> 
  filter(fin_mark == "AD") |> 
  count(fish_count) |>
  rename(frequency = n) |>
  ##fill in 0s
  full_join(expand_grid(fish_count = seq(0, max(dat.fit$fish_count), by = 1)
  )) |> 
  mutate(frequency = if_else(is.na(frequency),
                             0,
                             frequency)) |> 
  mutate(scenario = "observation")

gp.cur = bind_rows(pred.partial, obs.partial) |> 
  ggplot(aes(x = fish_count, y = frequency, fill = scenario))+
  geom_col(position = "dodge2", width = .6)+
  ggtitle("AD catch frequencies")+
  xlim(c(-1,21))+
  labs(subtitle = "using negbin model")+
  theme_bw(base_size = 14)

gp.cur


gp.cur = bind_rows(pred.partial, obs.partial) |> 
  filter(fish_count > 0) |> 
  ggplot(aes(x = fish_count, y = frequency, fill = scenario))+
  geom_col(position = "dodge2", width = .6)+
  ggtitle("AD catch frequencies (skipping 0s to provide more clarity)")+
  labs(subtitle = "using negbin model")+
  xlim(c(0,21))+
  theme_bw(base_size = 14)

gp.cur


## Unmarked

pred.partial = freq.pred |> 
  filter(fin_mark == "UM") |> 
  group_by(predicted_count) |> 
  summarize(expected_frequency = sum(expected_frequency)) |> 
  ungroup() |> 
  mutate(scenario = "prediction") |> 
  rename(fish_count = predicted_count, frequency = expected_frequency)

obs.partial = dat.fit |> 
  filter(fin_mark == "UM") |> 
  count(fish_count) |>
  rename(frequency = n) |>
  ##fill in 0s
  full_join(expand_grid(fish_count = seq(0, max(dat.fit$fish_count), by = 1)
  )) |> 
  mutate(frequency = if_else(is.na(frequency),
                             0,
                             frequency)) |> 
  mutate(scenario = "observation")

gp.cur = bind_rows(pred.partial, obs.partial) |> 
  ggplot(aes(x = fish_count, y = frequency, fill = scenario))+
  geom_col(position = "dodge2", width = .6)+
  ggtitle("UM catch frequencies")+
  xlim(c(-1,21))+
  labs(subtitle = "from `preliminary-catch-model.R`, using negbin model")+
  theme_bw(base_size = 14)

gp.cur

gp.cur = bind_rows(pred.partial, obs.partial) |> 
  filter(fish_count > 0) |> 
  ggplot(aes(x = fish_count, y = frequency, fill = scenario))+
  geom_col(position = "dodge2", width = .6)+
  xlim(c(0,21))+
  ggtitle("UM catch frequencies (skipping 0s to provide more clarity)")+
  labs(subtitle = "from `preliminary-catch-model.R`, using negbin model")+
  theme_bw(base_size = 14)

gp.cur

gp.cur = bind_rows(pred.partial, obs.partial) |> 
  filter(fish_count > 4) |> 
  ggplot(aes(x = fish_count, y = frequency, fill = scenario))+
  geom_col(position = "dodge2", width = .6)+
  xlim(c(4,21))+
  ggtitle("UM catch frequencies (skipping <5s to provide more clarity)")+
  labs(subtitle = "from `preliminary-catch-model.R`, using negbin model")+
  theme_bw(base_size = 14)

gp.cur


```

### a quick interpretation aside:

Does that one observation of 16 unmarked fish stand out? Well, the model thinks we have a small but decreasing probability of reports of more than 10 unmarked encounters. In total, how many people to we expect to have more than 10 unmarked fish?

```{r}
## quick thoughts:
freq.pred |> 
  filter(fin_mark == "UM") |> 
  filter(predicted_count > 10) |> 
  pull(expected_frequency) |> 
  sum()

```

About 3. So having one observation in that range is not surprising.

#### more aside

But the model thinks we should have at least 3 people reporting more than 10 fish, and we only had 1!

Isn't that a problem?

Not really. As we move far away from our actual data, I don't put high trust in the exact values of the model predictions. There's some math involved, but instead we can imagine that we're tiny model goblins in charge of determining the frequencies we should see for different fish counts. We know the frequencies should fall as we get to higher and higher fish counts, but how rapidly should they fall? Well, past 10 fish, we only have one actual data point, so we are mostly guessing based on the rate of fall in the smaller fish counts. If we get that rate of fall incorrect, the predicted frequencies of fish_counts in the 20s and 30s might be off, but as long as those predictions are small our data will still seem to be a pretty good match to the predictions.

As an example of this in practice, the model thinks there's a 10% chance of having one person someone reporting between 50 and 100 encountered across all the interviews in this dataset.

```{r}
foo <- function(x){dnbinom(50:100, size = cur.theta, mu = x)}
df = data.frame(mu = mus)|> 
  mutate(mat = do.call(rbind, purrr::map(mu, foo)))
sum(df$mat)
```

Biologically this seems pretty impossible, but this isn't really a failing of the model. Instead, the problem is that we're using model inferences far from values represented in the data we used to fit the model. We should avoid doing that.

### returning from the aised: how do our fish counts compare?

```{r}
cat("Total fish in our data")
sum(dat.fit$fish_count)
cat("Total fish predicted from our model")
sum(freq.pred$predicted_count * freq.pred$expected_frequency)
```

We're in the same ballpark, so probably fine.

### Rounding bias

```{r}
## Looking at rounding bias ===============================
obs.partial = dat.fit |> 
  count(fin_mark, fish_count) |> 
  rename(frequency = n) |> 
  filter(fish_count %in% c(5, 10, 15, 20)) |> 
  ## adding 0s where missing
  full_join(expand_grid(fin_mark = c("AD", "UM"), 
                        fish_count = c(5, 10, 15, 20))
  ) |> 
  mutate(frequency = if_else(is.na(frequency),
                             0,
                             frequency)) |> 
  mutate(scenario = "observation")

pred.partial = freq.pred |> 
  filter(predicted_count %in% c(5, 10, 15, 20)) |> 
  mutate(scenario = "prediction") |> 
  select(fin_mark, fish_count = predicted_count, frequency = expected_frequency, scenario)

bind_rows(obs.partial, pred.partial) |> 
  ggplot(aes(x = as.factor(fish_count), y = frequency, fill = scenario))+
  geom_col(position = "dodge2")+
  labs(x = "Fish reported",
       title = "Looking for rounding bias: comparing frequencies")+
  facet_wrap(.~ fin_mark, ncol = 1)+
  theme_bw(base_size = 14)

## let's look at predictions vs obs by category

obs.partial = dat.fit |> 
  count(fin_mark, round_cat1) |> 
  rename(frequency = n) |> 
  mutate(scenario = "observation") |> 
  na.omit()

pred.partial = freq.pred |> 
  group_by(fin_mark, round_cat1) |> 
  summarize(frequency = sum(expected_frequency)) |> 
  ungroup() |> 
  mutate(scenario = "prediction") |> 
  na.omit()

bind_rows(obs.partial, pred.partial) |> 
  ggplot(aes(x = round_cat1, y = frequency, fill = scenario))+
  geom_col(position = "dodge2")+
  labs(x = "Rounding category",
       title = "Looking for rounding bias: comparing within categories")+
  facet_wrap(.~ fin_mark, ncol = 1)+
  theme_bw(base_size = 14)

## let's look at predictions vs obs by category

obs.partial = dat.fit |> 
  count(fin_mark, round_cat2) |> 
  rename(frequency = n) |> 
  mutate(scenario = "observation") |> 
  na.omit()

pred.partial = freq.pred |> 
  group_by(fin_mark, round_cat2) |> 
  summarize(frequency = sum(expected_frequency)) |> 
  ungroup() |> 
  mutate(scenario = "prediction") |> 
  na.omit()

bind_rows(obs.partial, pred.partial) |> 
  ggplot(aes(x = round_cat2, y = frequency, fill = scenario))+
  geom_col(position = "dodge2")+
  labs(x = "Rounding category",
       title = "Looking for rounding bias: comparing within categories (2)")+
  facet_wrap(.~ fin_mark, ncol = 1)+
  theme_bw(base_size = 14)

```

### Prestige bias

```{r}

facet_names = c(`TRUE` = "Had kept fish",
                `FALSE` = "Did not have kept fish")

pred.partial = df.pred |> 
  filter(fin_mark == "UM") |> 
  group_by(predicted_count, had.kept) |> 
  summarize(expected_frequency = sum(predicted_probability)) |> 
  ungroup() |> 
  mutate(scenario = "prediction") |> 
  rename(fish_count = predicted_count, frequency = expected_frequency)

obs.partial = dat.fit |> 
  filter(fin_mark == "UM") |> 
  count(had.kept, fish_count) |>
  rename(frequency = n) |>
  ##fill in 0s
  full_join(expand_grid(had.kept = as.factor(c(TRUE, FALSE)), 
                        fish_count = seq(0, max(dat.fit$fish_count), by = 1)
  )) |> 
  mutate(frequency = if_else(is.na(frequency),
                             0,
                             frequency)) |> 
  mutate(scenario = "observation")


bind_rows(pred.partial, obs.partial) |> 
  # filter(fish_count > 0) |>
  ggplot(aes(x = fish_count, y = frequency, fill = scenario))+
  geom_col(position = "dodge2", width = .6)+
  facet_wrap(.~ had.kept, labeller = as_labeller(facet_names),
             ncol = 1, scale = "free_y") +
  xlim(c(-1, 20))+
  ggtitle("UM catch frequencies")+
  labs(subtitle = "excluding 0s for visual clarity")+
  theme_bw(base_size = 14)

bind_rows(pred.partial, obs.partial) |>
  filter(fish_count > 0) |>
  ggplot(aes(x = fish_count, y = frequency, fill = scenario))+
  geom_col(position = "dodge2", width = .6)+
  facet_wrap(.~ had.kept, labeller = as_labeller(facet_names),
             ncol = 1, scale = "free_y") +
  xlim(c(NA, 20))+
  ggtitle("UM catch frequencies")+
  labs(subtitle = "excluding 0s for visual clarity")+
  theme_bw(base_size = 14)


bind_rows(pred.partial, obs.partial) |>
  filter(fish_count > 4) |>
  ggplot(aes(x = fish_count, y = frequency, fill = scenario))+
  geom_col(position = "dodge2", width = .6)+
  facet_wrap(.~ had.kept, labeller = as_labeller(facet_names),
             ncol = 1, scale = "free_y") +
  xlim(c(NA, 20))+
  ggtitle("UM catch frequencies")+
  labs(subtitle = "excluding <5s for visual clarity")+
  theme_bw(base_size = 14)
```

That does not look like the high residuals were associated just with "did not have kept fish". In fact, the discrepency between prediction and obs seems even larger for the "had kept fish" category.

:::

## Zero-inflated model

Our model above looks like it may have had dispersion issues. Note that it is important that we try to find a model that correctly represents dispersion, as we're trying to look at the mismatch between observations and expectations for a range of fish_counts, including large fish_counts.

### Model comparisons

If our model is too constrained in terms of dispersion, we may be "forcing" discrepencies. A few ways to address dispersion:

-   Are we zero-inflated? Tests above did not seem to indicate that we were, but this is still worth checking
-   Is dispersion different between marked and unmarked fish?
-   Should we be using a different version of the negative binomial? e.g. scaling quadratically rather than linearly with the sum of our predictors?

Here I test a series of model structures to see what has the best support using AIC (for AIC, the lowest is the best.

```{r}

## Zero-inflated model ----------------------------------
# Looking at the predicted vs observed distributions of UM and AD, it looks like we're running into differing dispersion. 
# I think the best method is to try to account for this with a zero-inflated model


# Let's check out the nbinom2 family in glmmTMB
#
library(glmmTMB)

df.models = NULL

out.tmba = glmmTMB(fish_count ~ s(doy, fin_mark, bs = 'fs', k = 20) + fishing_duration_minutes + (1 | yearfac),
                   family = nbinom1,
                   REML = TRUE,
                   data = dat.fit)

df.models = bind_rows(df.models,
                      data.frame(name = "out.tmba", 
                                 aic = AIC(out.tmba),
                                 family = family(out.tmba)[[1]],
                                 explanation = "null model. Our 'simple gam model' above, but fit using glmmTMB")
)

out.tmbb = glmmTMB(fish_count ~ s(doy, fin_mark, bs = 'fs', k = 20) + fishing_duration_minutes + (1 | yearfac),
                   family = nbinom1,
                   dispformula = ~fin_mark,
                   REML = TRUE,
                   data = dat.fit)

df.models = bind_rows(df.models,
                      data.frame(name = "out.tmbb", 
                                 aic = AIC(out.tmbb),
                                 family = family(out.tmbb)[[1]],
                                 explanation = "Null model, but with different dispersions based on fin mark"))

out.tmb1 = glmmTMB(fish_count ~ s(doy, fin_mark, bs = 'fs', k = 20) + fishing_duration_minutes + (1 | yearfac),
                   ziformula = ~ fin_mark,
                   family = nbinom1,
                   REML = TRUE,
                   data = dat.fit)

df.models = bind_rows(df.models,
                      data.frame(name = "out.tmb1", 
                                 aic = AIC(out.tmb1),
                                 family = family(out.tmb1)[[1]],
                                 explanation = "zero inflated model with zinf ~ fin_mark, no other dispersion modifications"))

out.tmb2 = glmmTMB(fish_count ~ s(doy, fin_mark, bs = 'fs', k = 20) + fishing_duration_minutes + (1 | yearfac),
                   ziformula = ~ fin_mark,
                   family = nbinom2,
                   REML = TRUE,
                   data = dat.fit)

df.models = bind_rows(df.models,
                      data.frame(name = "out.tmb2", 
                                 aic = AIC(out.tmb2),
                                 family = family(out.tmb2)[[1]],
                                 explanation = "zero inflated model with zinf ~ fin_mark, using the nbinom2 family (quadratic)"))

out.tmb3 = glmmTMB(fish_count ~ s(doy, fin_mark, bs = 'fs', k = 20) + fishing_duration_minutes + (1 | yearfac),
                   ziformula = ~ fin_mark,
                   family = nbinom1,
                   dispformula = ~fin_mark,
                   REML = TRUE,
                   data = dat.fit)


df.models = bind_rows(df.models,
                      data.frame(name = "out.tmb3", 
                                 aic = AIC(out.tmb3),
                                 family = family(out.tmb3)[[1]],
                                 explanation = "zero inflated model with zinf ~ fin_mark, dispersion/shape term varies by fin mark"))

out.tmb4 = glmmTMB(fish_count ~ s(doy, fin_mark, bs = 'fs', k = 20) + fishing_duration_minutes + (1 | yearfac),
                   ziformula = ~ fin_mark,
                   family = truncated_nbinom1,
                   dispformula = ~fin_mark,
                   REML = TRUE,
                   data = dat.fit)

df.models = bind_rows(df.models,
                      data.frame(name = "out.tmb4", 
                                 aic = AIC(out.tmb4),
                                 family = family(out.tmb4)[[1]],
                                 explanation = "hurdle model with zinf ~ fin_mark, dispersion/shape term varies by fin mark"))

out.tmb4b = glmmTMB(fish_count ~ s(doy, fin_mark, bs = 'fs', k = 20) + fishing_duration_minutes + (1 | yearfac),
                    ziformula = ~ fin_mark,
                    family = truncated_nbinom1,
                    REML = TRUE,
                    data = dat.fit)

df.models = bind_rows(df.models,
                      data.frame(name = "out.tmb4b", 
                                 aic = AIC(out.tmb4b),
                                 family = family(out.tmb4b)[[1]],
                                 explanation = "hurdle model with zinf ~ fin_mark, no other dispersion modifications"))



## more overdispersion?
out.tmb3b = glmmTMB(fish_count ~ s(doy, fin_mark, bs = 'fs') + fishing_duration_minutes + (1 | yearfac) + (1 | interview_id),
                    ziformula = ~ fin_mark,
                    family = nbinom1,
                    dispformula = ~fin_mark,
                    REML = TRUE,
                    data = dat.fit)

df.models = bind_rows(df.models,
                      data.frame(name = "out.tmb3b", 
                                 aic = AIC(out.tmb3b),
                                 family = family(out.tmb3b)[[1]],
                                 explanation = "zero inflated model with zinf ~ fin_mark, dispersion/shape term varies by fin mark, also individual-based random effect"))


## more overdispersion?
out.tmb4 = glmmTMB(fish_count ~ s(doy, fin_mark, bs = 'fs', k = 20) + fishing_duration_minutes + (1 | yearfac),
                   ziformula = ~ .,
                   family = nbinom1,
                   dispformula = ~fin_mark,
                   REML = TRUE,
                   data = dat.fit)

df.models = bind_rows(df.models,
                      data.frame(name = "out.tmb4", 
                                 aic = AIC(out.tmb4),
                                 family = family(out.tmb4)[[1]],
                                 explanation = "zero inflated model with zinf ~ [model structure], dispersion/shape term varies by fin mark"))

df.models |> 
  arrange(aic) |> 
  gt()
```

We see the most support for `out.tmb3b` (sorry for the terrible naming conventions). This model has (a) zero-inflation that depends on fin marks, (b) separate dispersion/shape parameter estimation for marked and unmarked fish, and (c) an individual-based random effect. (c) is an especially confusing concept, but allows for additional overdispersion. See \[I have a ref for this on my personal computer, need to grab it\].

We will move forward using this best model.

::: panel-tabset
### Model diagnostics

```{r}
## rename for simplicity
out.tmb = out.tmb3b
resids = simulateResiduals(out.tmb)
plot(resids)
```

This looks much better than our gam diagnostics.

### Calculating predictions

```{r}
mus = predict(out.tmb, newdata = dat.fit, type = "conditional")
probs.struct.zero = predict(out.tmb, newdata = dat.fit, type = "zprob") 
dispersion.vals = predict(out.tmb, newdata = dat.fit, type = "disp") 
#For nbinom1: V=\mu(1+\phi)
#For nbinom2:  V=\mu(1+\mu/\phi) = \mu+\mu^2/\phi

cli::cli_alert("It appears that the dispersional value as pulled out of the model with sigma() is equivalent to the size parameter of the dnbinom function. This is worth checking")



foo <- function(mu, disp.val, zprob){
  non.struct = dnbinom(0:30, size = disp.val, mu = mu)
  struct = c(1, rep(0,30))
  return(non.struct * (1 - zprob) + struct * zprob)
}
## this gets messy



df = data.frame(mu = mus,
                disp.val = dispersion.vals,
                probs.struct.zero
)|> 
  mutate(mat = do.call(rbind, purrr::pmap(
    list(mu = mu, disp.val = disp.val, zprob = probs.struct.zero),
    foo)))

df.pred = as.data.frame(df$mat)
#gross names for a sec:
names(df.pred) = 0:30  
## now combine with the data and use pivoting to put into handy form
df.pred <- dat.fit |> 
  select(-fish_count_predicted, -fish_count_predicted_int) |> 
  cbind(df.pred) |> 
  pivot_longer(cols = "0":"30",
               names_to = "predicted_count",
               values_to = "predicted_probability") |> 
  mutate(predicted_count = as.numeric(predicted_count))

## With this dataframe, we can now sum up the number of counts we would expect to see
## for marked and unmarked fish, *accounting* for the variability in negative binomials

freq.pred = df.pred |> 
  group_by(fin_mark, predicted_count) |> 
  summarize(expected_frequency = sum(predicted_probability)) |> 
  ungroup() |> 
  ## adding rounding categorizations
  mutate(round_cat1 = case_when(
    predicted_count >= 3 & predicted_count <= 7 ~ "3-7",
    predicted_count >= 8 & predicted_count <= 12 ~ "8-12",
    predicted_count >= 13 ~ "13+",
  )) |> 
  mutate(round_cat1 = factor(round_cat1, levels = c("3-7", "8-12", "13+"))) |> 
  mutate(round_cat2 = case_when(
    predicted_count >= 3 & predicted_count <= 7 ~ "3-7",
    predicted_count >= 8 ~ "8+"
  )) |> 
  mutate(round_cat2 = factor(round_cat2, levels = c("3-7", "8+")))

## basic plotting ================================

pred.partial = freq.pred |> 
  filter(fin_mark == "AD") |> 
  group_by(predicted_count) |> 
  summarize(expected_frequency = sum(expected_frequency)) |> 
  ungroup() |> 
  mutate(scenario = "prediction") |> 
  rename(fish_count = predicted_count, frequency = expected_frequency)

obs.partial = dat.fit |> 
  filter(fin_mark == "AD") |> 
  count(fish_count) |>
  rename(frequency = n) |>
  ##fill in 0s
  full_join(expand_grid(fish_count = seq(0, max(dat.fit$fish_count), by = 1)
  )) |> 
  mutate(frequency = if_else(is.na(frequency),
                             0,
                             frequency)) |> 
  mutate(scenario = "observation")

gp.cur = bind_rows(pred.partial, obs.partial) |> 
  ggplot(aes(x = fish_count, y = frequency, fill = scenario))+
  geom_col(position = "dodge2", width = .6)+
  ggtitle("AD catch frequencies")+
  labs(subtitle = "using zero-inflated model")+
  theme_bw(base_size = 14)

gp.cur

gp.cur = bind_rows(pred.partial, obs.partial) |> 
  filter(fish_count > 0) |> 
  ggplot(aes(x = fish_count, y = frequency, fill = scenario))+
  geom_col(position = "dodge2", width = .6)+
  ggtitle("AD catch frequencies (skipping 0s to provide more clarity)")+
  labs(subtitle = "using zero-inflated model")+
  theme_bw(base_size = 14)

gp.cur

pred.partial = freq.pred |> 
  filter(fin_mark == "UM") |> 
  group_by(predicted_count) |> 
  summarize(expected_frequency = sum(expected_frequency)) |> 
  ungroup() |> 
  mutate(scenario = "prediction") |> 
  rename(fish_count = predicted_count, frequency = expected_frequency)

obs.partial = dat.fit |> 
  filter(fin_mark == "UM") |> 
  count(fish_count) |>
  rename(frequency = n) |>
  ##fill in 0s
  full_join(expand_grid(fish_count = seq(0, max(dat.fit$fish_count), by = 1)
  )) |> 
  mutate(frequency = if_else(is.na(frequency),
                             0,
                             frequency)) |> 
  mutate(scenario = "observation")

gp.cur = bind_rows(pred.partial, obs.partial) |> 
  ggplot(aes(x = fish_count, y = frequency, fill = scenario))+
  geom_col(position = "dodge2", width = .6)+
  ggtitle("UM catch frequencies")+
  labs(subtitle = "using zero-inflated model")+
  theme_bw(base_size = 14)

gp.cur

gp.cur = bind_rows(pred.partial, obs.partial) |> 
  filter(fish_count > 0) |> 
  ggplot(aes(x = fish_count, y = frequency, fill = scenario))+
  geom_col(position = "dodge2", width = .6)+
  ggtitle("UM catch frequencies (skipping 0s to provide more clarity)")+
  labs(subtitle = "using negbin model")+
  theme_bw(base_size = 14)

gp.cur


gp.cur = bind_rows(pred.partial, obs.partial) |> 
  filter(fish_count > 4) |> 
  ggplot(aes(x = fish_count, y = frequency, fill = scenario))+
  geom_col(position = "dodge2", width = .6)+
  ggtitle("UM catch frequencies (skipping <5s to provide more clarity)")+
  labs(subtitle = "using negbin model")+
  theme_bw(base_size = 14)

gp.cur
```

### Rounding bias

```{r}
## Looking at rounding bias ===============================
obs.partial = dat.fit |> 
  count(fin_mark, fish_count) |> 
  rename(frequency = n) |> 
  filter(fish_count %in% c(5, 10, 15, 20)) |> 
  ## adding 0s where missing
  full_join(expand_grid(fin_mark = c("AD", "UM"), 
                        fish_count = c(5, 10, 15, 20))
  ) |> 
  mutate(frequency = if_else(is.na(frequency),
                             0,
                             frequency)) |> 
  mutate(scenario = "observation")

pred.partial = freq.pred |> 
  filter(predicted_count %in% c(5, 10, 15, 20)) |> 
  mutate(scenario = "prediction") |> 
  select(fin_mark, fish_count = predicted_count, frequency = expected_frequency, scenario)

bind_rows(obs.partial, pred.partial) |> 
  ggplot(aes(x = as.factor(fish_count), y = frequency, fill = scenario))+
  geom_col(position = "dodge2")+
  labs(x = "Fish reported",
       title = "Looking for rounding bias: comparing frequencies")+
  facet_wrap(.~ fin_mark, ncol = 1)+
  theme_bw(base_size = 14)

## let's look at predictions vs obs by category

obs.partial = dat.fit |> 
  count(fin_mark, round_cat1) |> 
  rename(frequency = n) |> 
  mutate(scenario = "observation") |> 
  na.omit()

pred.partial = freq.pred |> 
  group_by(fin_mark, round_cat1) |> 
  summarize(frequency = sum(expected_frequency)) |> 
  ungroup() |> 
  mutate(scenario = "prediction") |> 
  na.omit()

bind_rows(obs.partial, pred.partial) |> 
  ggplot(aes(x = round_cat1, y = frequency, fill = scenario))+
  geom_col(position = "dodge2")+
  labs(x = "Rounding category",
       title = "Looking for rounding bias: comparing within categories")+
  facet_wrap(.~ fin_mark, ncol = 1)+
  theme_bw(base_size = 14)

## let's look at predictions vs obs by category

obs.partial = dat.fit |> 
  count(fin_mark, round_cat2) |> 
  rename(frequency = n) |> 
  mutate(scenario = "observation") |> 
  na.omit()

pred.partial = freq.pred |> 
  group_by(fin_mark, round_cat2) |> 
  summarize(frequency = sum(expected_frequency)) |> 
  ungroup() |> 
  mutate(scenario = "prediction") |> 
  na.omit()

bind_rows(obs.partial, pred.partial) |> 
  ggplot(aes(x = round_cat2, y = frequency, fill = scenario))+
  geom_col(position = "dodge2")+
  labs(x = "Rounding category",
       title = "Looking for rounding bias: comparing within categories (2)")+
  facet_wrap(.~ fin_mark, ncol = 1)+
  theme_bw(base_size = 14)

```

### Prestige bias

```{r}

facet_names = c(`TRUE` = "Had kept fish",
                `FALSE` = "Did not have kept fish")

pred.partial = df.pred |> 
  filter(fin_mark == "UM") |> 
  group_by(predicted_count, had.kept) |> 
  summarize(expected_frequency = sum(predicted_probability)) |> 
  ungroup() |> 
  mutate(scenario = "prediction") |> 
  rename(fish_count = predicted_count, frequency = expected_frequency)

obs.partial = dat.fit |> 
  filter(fin_mark == "UM") |> 
  count(had.kept, fish_count) |>
  rename(frequency = n) |>
  ##fill in 0s
  full_join(expand_grid(had.kept = as.factor(c(TRUE, FALSE)), 
                        fish_count = seq(0, max(dat.fit$fish_count), by = 1)
  )) |> 
  mutate(frequency = if_else(is.na(frequency),
                             0,
                             frequency)) |> 
  mutate(scenario = "observation")


bind_rows(pred.partial, obs.partial) |> 
  # filter(fish_count > 0) |>
  ggplot(aes(x = fish_count, y = frequency, fill = scenario))+
  geom_col(position = "dodge2", width = .6)+
  facet_wrap(.~ had.kept, labeller = as_labeller(facet_names),
             ncol = 1, scale = "free_y") +
  xlim(c(-1, 20))+
  ggtitle("UM catch frequencies")+
  labs(subtitle = "excluding 0s for visual clarity")+
  theme_bw(base_size = 14)

bind_rows(pred.partial, obs.partial) |>
  filter(fish_count > 0) |>
  ggplot(aes(x = fish_count, y = frequency, fill = scenario))+
  geom_col(position = "dodge2", width = .6)+
  facet_wrap(.~ had.kept, labeller = as_labeller(facet_names),
             ncol = 1, scale = "free_y") +
  xlim(c(NA, 20))+
  ggtitle("UM catch frequencies")+
  labs(subtitle = "excluding 0s for visual clarity")+
  theme_bw(base_size = 14)


bind_rows(pred.partial, obs.partial) |>
  filter(fish_count > 4) |>
  ggplot(aes(x = fish_count, y = frequency, fill = scenario))+
  geom_col(position = "dodge2", width = .6)+
  facet_wrap(.~ had.kept, labeller = as_labeller(facet_names),
             ncol = 1, scale = "free_y") +
  xlim(c(NA, 20))+
  ggtitle("UM catch frequencies")+
  labs(subtitle = "excluding <5s for visual clarity")+
  theme_bw(base_size = 14)
```
:::
:::::
