---
title: "Creel survey sampling evaluation"
execute:
    echo: false
    cache: true
author: "Evan Booher"
format:
  docx: 
    reference-doc: "WDFW_Report_Template.docx"
    fig-dpi: 600
editor: visual
params:
  focal_waterbody: "Nisqually River"
  study_design: "Standard"  
  boat_type_collapse: "Yes"
  fish_location_determines_type: "No"
  angler_type_kayak_pontoon: "bank" 
  person_count_type: "angler"
  period_pe: "week"
  period_bss: "day"
  days_wkend: "!r c('Friday','Saturday', 'Sunday')"
  day_length_expansion: "night closure"
  min_fishing_time: 0.5
---

```{r libraries}
options(scipen = 999)

library(CreelEstimateR)
library(creelutils)
library(tidyverse)
library(stringr)
library(here)
library(gt)
library(glue)
library(tidymodels)
library(kableExtra)
library(flextable)
library(patchwork)
library(gridExtra)
library(knitr)
# library(kableExtra)
# library(flextable)
```

```{r local edited function for processing interview data}
#| cache: false
# local edited function from creel_estimates 
source(here::here("scripts", "prep_dwg_interview_angler_types_1.1.R"))   
```


```{r}
#| include: false

#lists that contain creel datasets downloaded from data.wa.gov
dwg <- creelutils::get_fishery_data(fishery_names = "Nisqually salmon", year = 2021:2023)

# add date grouping variables
dwg$interview <- dwg$interview |> 
  mutate(
    month = lubridate::month(.data$event_date),
    year = lubridate::year(.data$event_date),
    week = lubridate::week(.data$event_date)
  )

# add date grouping variables
dwg$effort <- dwg$effort |>
  mutate(
    month = lubridate::month(.data$event_date),
    year = lubridate::year(.data$event_date),
    week = lubridate::week(.data$event_date)
  ) 

eff21 <- dwg$effort |>
  filter(year == 2021) |> 
   mutate(
    # extract one or more digits at end of comments, then convert to integer
    count_sequence = str_extract(comments, "\\d+$") %>% 
                     as.integer()
  )

dwg$effort  <- dwg$effort |> 
  filter(year %in% c(2022,2023)) |> 
  rbind(eff21)

```

```{r links to data, single fishery}
#| eval: false
#| results: asis

fishery_name <- "Nisqually salmon 2021"

dwg_base <- list(effort = "https://data.wa.gov/resource/h9a6-g38s.csv", 
    interview = "https://data.wa.gov/resource/rpax-ahqm.csv", 
    catch = "https://data.wa.gov/resource/6y4e-8ftk.csv", 
    water_bodies = "https://data.wa.gov/resource/nbd2-vdmz.csv", 
    closures = "https://data.wa.gov/resource/6zm6-iep6.csv", 
    fishery_manager = "https://data.wa.gov/resource/vkjc-s5u8.csv")


# Function to encode URL
encode_url <- function(base_url, fishery_name) {
  utils::URLencode(paste0(base_url, "?$where=fishery_name='", fishery_name, "'&$limit=100000"))
}

# Apply function to all base URLs
links <- map(dwg_base, encode_url, fishery_name)


# Generate Markdown links using the names of the endpoints for the link text.
markdown_links <- map2_chr(names(links), links, 
  ~ paste0("[", .x, "](", .y, ")")
)

# Output the Markdown links.
# markdown_links

# Create a data frame for the table
df <- data.frame(
  Link = markdown_links,
  stringsAsFactors = FALSE
)

# Render a table where the markdown in the "Link" column is not escaped.
# When you output this in a chunk with results = 'asis', Quarto will convert these into clickable links in Word.
kable(df, escape = FALSE)

# 
# # Generate markdown links
# map_chr(links, ~paste0("[Download public data](", .x, ")"))


# links$effort <- utils::URLencode(paste0(dwg_base$effort,"?$where=fishery_name='", fishery_name, "'&$limit=100000"))
# links$interview <- utils::URLencode(paste0(dwg_base$interview, "?$where=fishery_name='", fishery_name, "'&$limit=100000"))
# links$catch <- utils::URLencode(paste0(dwg_base$catch, "?$where=fishery_name='", fishery_name, "'&$limit=100000"))
# links$closures <- utils::URLencode(paste0(dwg_base$closures, "?$where=fishery_name='", fishery_name, "'&$limit=100000"))
# links$locations <- utils::URLencode(paste0(dwg_base$fishery_manager, "?$where=fishery_name='", fishery_name, "'&$limit=100000"))
# 
# paste0("[Download public data](", links$effort , ")")
# paste0("[Download public data](", links$interview , ")")
# paste0("[Download public data](", links$catch , ")")
# paste0("[Download public data](", links$closures , ")")
# paste0("[Download public data](", links$locations , ")")
# paste0("[Download public data](", links$effort , ")")
```

```{r generate data links for multiple fisheries}
library(knitr)
#| results: asis 
#| eval: false

# Define a vector of fishery names.
fishery_names <- dwg$interview |> distinct(fishery_name) |> pull()

# Define your base endpoints.
dwg_base <- list(
  effort          = "https://data.wa.gov/resource/h9a6-g38s.csv", 
  interview       = "https://data.wa.gov/resource/rpax-ahqm.csv", 
  catch           = "https://data.wa.gov/resource/6y4e-8ftk.csv", 
  water_bodies    = "https://data.wa.gov/resource/nbd2-vdmz.csv", 
  closures        = "https://data.wa.gov/resource/6zm6-iep6.csv", 
  fishery_manager = "https://data.wa.gov/resource/vkjc-s5u8.csv"
)

# Function to build and URL-encode the full query URL for a given fishery.
encode_url <- function(base_url, fishery_name) {
  utils::URLencode(paste0(base_url, "?$where=fishery_name='", fishery_name, "'&$limit=100000"))
}

# Create a matrix (later converted to a data frame) with one column per fishery and one row per endpoint.
# Each cell gets a clickable Markdown link whose text is the endpoint name.
result <- sapply(fishery_names, function(fishery) {
  sapply(names(dwg_base), function(endpoint) {
    url <- encode_url(dwg_base[[endpoint]], fishery)
    paste0("[", endpoint, "](", url, ")")
  })
}, simplify = TRUE, USE.NAMES = TRUE)

# Convert the result to a data frame with row names corresponding to each data component.
result_df <- as.data.frame(result, stringsAsFactors = FALSE, row.names = FALSE)
# rownames(result_df) <- names(dwg_base)

# Render the table.
# Use results='asis' in your Quarto R chunk to output this table as raw Markdown,
# so Pandoc interprets the Markdown links correctly when converting to Word.
kable(result_df, escape = FALSE, align = "l", caption = "Table _. Links to creel survey data available through the state of Washington's open data portal, https://data.wa.gov.")


```

```{r fisheries_list}

fisheries <- c(
  "Nisqually salmon 2022",
  "Nisqually salmon 2023"
)


all_fisheries <- c(fisheries)

```

```{r query data}

# fisheries <- c("Skagit fall salmon 2024")

catch <- list()
effort <- list()

# Fetch / load data
# Prompts for user password if ran interactively and actively queries the creel database. This result is saved locally.
# However, if rendering the quarto document, it will load the local copy of the model estimates.
if (interactive()) {
  creel_estimates <- list()
  
  # establish connection
  con <- creelutils::establish_db_con()

  # query analysis_lut
  creel_estimates$analysis_lut <- creelutils::fetch_db_table(
    con, "creel", "model_analysis_lut") |> 
    # fishery name ends at the first underscore of analysis name field
    mutate(fishery_name = stringr::str_extract(analysis_name, "^[^_]+")) |> 
    relocate(fishery_name, .after = "analysis_name") |> 
    filter(fishery_name %in% all_fisheries)
  
  # query total table
  # creel_estimates$total <- creelutils::fetch_db_table(
  #   con, "creel", "vw_model_estimates_total") |> 
  #   filter(analysis_id %in% c("dbca0cf7-ff69-4c52-ab05-73d917b8c5aa", "80e4bde0-1795-467c-8c86-b820d6edd896"))
  
  creel_estimates$total <- creelutils::fetch_db_table(
    con, "creel", "vw_model_estimates_total", filter = c("analysis_id %in% c('dbca0cf7-ff69-4c52-ab05-73d917b8c5aa', '80e4bde0-1795-467c-8c86-b820d6edd896')"))
  
  # query stratum table
  creel_estimates$stratum <- creelutils::fetch_db_table(
    con, "creel", "vw_model_estimates_stratum", filter = c("analysis_id %in% c('dbca0cf7-ff69-4c52-ab05-73d917b8c5aa', '80e4bde0-1795-467c-8c86-b820d6edd896')")) 
# |> 
    # filter(analysis_id %in% c("dbca0cf7-ff69-4c52-ab05-73d917b8c5aa", "80e4bde0-1795-467c-8c86-b820d6edd896"))
  
  # save queried model estimates as local file
  # needed for knitting as rstudioapi:: functions cannot be used during a Quarto render
  save(creel_estimates, file = "./creel_estimates.Rdata")
  
} else {
  # load previous export of estimates saved locally during a Quarto render
  load("./creel_estimates.Rdata")
}

# retrieve the needed analysis id's 

# analysisid <- creel_estimates$analysis_lut |> slice_max(created_datetime) |> pull(analysis_id)

# add date variables 
creel_estimates$stratum <- creel_estimates$stratum |> 
     mutate(
    month = lubridate::month(.data$min_event_date),
    year = lubridate::year(.data$min_event_date),
    week = lubridate::week(.data$min_event_date),
    julian_date = lubridate::yday(.data$min_event_date)
  )


```

```{r report_tables}

# empty list of objects for report 
report_tabs <- list()

```

```{r read in all fishery dates}
# survey frame representing all days open and closed in each fishery year created these objects in separate instances of CreelEstimates code using prep_days() and export them to data folder HumptulipsCreel repo
# Specify file paths
file_paths <- c(
  here("cleaned_data", "days_Nisqually_2021.Rds"),
  here("cleaned_data", "days_Nisqually_2022.Rds"),
  here("cleaned_data", "days_Nisqually_2023.Rds"))

report_tabs$all_dates <- file_paths |> 
  map(read_rds) |> 
  bind_rows() |> 
  mutate(
    open_section_1 = coalesce(open_section_1, open_section_2),
    open_section_3 = coalesce(open_section_3, open_section_2),
    open_section_4 = coalesce(open_section_4, open_section_2)
  )


```

```{r survey day evaluation assuming section 2 represents sampling effort for entire fishery}
#| eval: true
#| include: true

# data wrangling to evaluate survey coverage - version aggregating over sections. If closure structure differed by section, do not use this code chunk, see chunk below including all sections

# use the effort data as the basis to determine if a creel survey occurred. This is imperfect and could be improved with a more exhaustive "check" for criteria across event, effort, and interview data components to determine if a creel survey was completed
report_tabs$survey_dates <- dwg$effort |> 
  filter(tie_in_indicator == 0, section_num == 2) |>  
  distinct(event_date) |> 
  mutate(
    surveyed = TRUE
  )

# perform an expanding left join of the full survey frame represented by the all_dates object to the survey_dates object, pivot long the open and surveyed status columns with True/False values 
report_tabs$combined_dates <- report_tabs$all_dates |> 
  left_join(report_tabs$survey_dates, by = "event_date") |> 
   mutate(across(
    .cols = contains("surveyed"),
    .fns = ~ replace_na(.x, FALSE)
  )) |> 
  select(-c(open_section_1, open_section_3, open_section_4)) |> 
  rename(open_section = open_section_2) |> # Nisqually, section 2 was consistently surveyed across all years
  select(event_date, starts_with(c("open", "surveyed"))) |> 
  pivot_longer(cols = starts_with(c("open", "surveyed")), 
               names_to = c(".value")) |> 
  mutate(year = year(event_date), month = month(event_date)) 
# |> 
#   mutate( # Nisqually specific add on 
#     monitoring_period = if_else(event_date >= "2021-08-02", TRUE, FALSE)
#   )

# Across the entire monitoring period, evaluate the number of open fishing days, survey days, proportion of open fishing days surveyed, and the number of closure days

# report_tabs$survey_dates_summary <- report_tabs$combined_dates |>
#   summarise(
#     survey_days = sum(across(starts_with("surveyed"), ~ .x), na.rm = TRUE),
#     closure_days = sum(across(starts_with("open"), ~ !.x), na.rm = TRUE),
#     total_fishing_days = n() - closure_days,
#     proportion_surveyed = round(survey_days/total_fishing_days, 2)
#   )

report_tabs$survey_dates_summary <- report_tabs$combined_dates |> 
  summarise(
    total_days = n(),
    survey_days = sum(across(starts_with("surveyed"), ~ .x), na.rm = TRUE),
    closure_days = sum(across(starts_with("open"), ~ !.x), na.rm = TRUE),
    # monitoring_period = sum(across(starts_with("monitoring"), ~ .x), na.rm = TRUE),
    total_fishing_days = n() - closure_days,
    # total_fishing_days_monitoring_period = monitoring_period - closure_days,
    proportion_surveyed_all_dates = round(survey_days/total_fishing_days, 2)
    # ,
    # proportion_surveyed_monitoring_period = round(survey_days/total_fishing_days_monitoring_period, 2)
  )

# Grouping for year and month, evaluate the number of open fishing days, survey days, proportion of open fishing days surveyed, and the number of closure days
report_tabs$combined_dates_monthly_p <- report_tabs$combined_dates |> 
  select(event_date, open_section
         # , monitoring_period
         ) |> # assumes by default single section section_num = 1
  left_join(report_tabs$survey_dates, by = "event_date") |> 
   mutate(across(
    .cols = contains("surveyed"),
    .fns = ~ replace_na(.x, FALSE)
  )) |> 
  select(event_date, starts_with(c("open", "surveyed"
                                   # , "monitoring"
                                   ))) |> 
  mutate(year = year(event_date), month = month(event_date)) |> 
  group_by(year, month) |>
summarise(
    survey_days = sum(across(starts_with("surveyed"), ~ .x), na.rm = TRUE),
    closure_days = sum(across(starts_with("open"), ~ !.x), na.rm = TRUE),
    # monitoring_period = sum(across(starts_with("monitoring"), ~ .x), na.rm = TRUE),
    total_fishing_days = n() - closure_days,
    # total_fishing_days_monitoring_period = monitoring_period - closure_days,
    proportion_surveyed_all_dates = round(survey_days/total_fishing_days, 2)
  ) 

```

# Results

## Creel survey sampling evaluation

### Survey days and closures

During the monitoring period from 2021 to 2023, a roving creel survey was conducted on the `r params$focal_waterbody` on `r report_tabs$survey_dates_summary[2]` out of the `r report_tabs$survey_dates_summary[1]` (`r round((report_tabs$survey_dates_summary[5]*100), 0)`%) days that the fishery was open.

```{r fishing days table}
#| eval: false
#| include: true

# At the monthly time step the mean sampled proportion of open fishing days was `r report_tabs$survey_dates_monthly_summary$mean`

report_tabs$combined_dates_monthly_p |>
  select(month, open_fishing_days = total_fishing_days, survey_days, closure_days) |>
  rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence()) |>
   knitr::kable(caption = "Table _. The number of open fishing days, survey days, proportion of open fishing days within each monitoring period surveyed, and the number of closure days per month during the Nisqually River salmon fishery from 2021 to 2023.")

```

```{r fishing days plot}
#| fig.cap: "Figure _. The number of open fishing days, survey days, and closure days per month during the Nisqually River salmon fishery from 2021 to 2023."
#| fig-width: 8
#| 
fishing_days <- report_tabs$combined_dates_monthly_p |>
  dplyr::select(year, month, open_fishing_days = total_fishing_days, closure_days, survey_days) |> 
  pivot_longer(cols = c(open_fishing_days, survey_days, closure_days),
               names_to = "day_type",
               values_to = "days") |> 
  mutate(day_type = factor(day_type, levels = c("open_fishing_days", "survey_days", "closure_days", "total_days")))
  
fishing_days |> 
  ggplot(aes(x = as.factor(month), y = days, fill = day_type)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  facet_wrap(~ year, scales = "free_x") +
  labs(x = "Month",
       y = "Number of days",
       fill = "Day category") +
  scale_fill_manual(values = c("open_fishing_days" = "steelblue", "survey_days" = "orange", "closure_days" = "red"),
                    labels = c("open_fishing_days" = "Open fishing days", "survey_days" = "Survey days", "closure_days" = "Closure days")
                    ) +
  theme_bw()

```

```{r proportion fishing days plot}
#| fig.cap: "Figure _. The proportion of days open to fishing per month that a roving creel survey was conducted during the Nisqually River salmon fishery from 2021 to 2023."

report_tabs$combined_dates_monthly_p |>
# drop_na(proportion_surveyed_monitoring_period) |> 
ggplot(aes(x = month, y = proportion_surveyed_all_dates)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  facet_wrap(~ year, ncol = 1, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
       x = "Month",
       y = "Proportion of open fishing days surveyed") +
  theme_bw()

```

```{r interview fishing time / angler type assigments}
#| cache: false

# use prep_dwg_interview_fishing_time, prep_dwg_interview_angler_types functions to calculate fishing time and create angler type assignments 
dwg$interview <- 
    CreelEstimateR::prep_dwg_interview_fishing_time(
      dwg_interview = dwg$interview,
      person_count_type = params$person_count_type,
      min_fishing_time = params$min_fishing_time,
      study_design = params$study_design
    )

dwg$interview <- 
    prep_dwg_interview_angler_types(
      interview_fishing_time = dwg$interview,
      study_design = params$study_design,
      boat_type_collapse = params$boat_type_collapse,
      fish_location_determines_type = params$fish_location_determines_type,
      angler_type_kayak_pontoon = params$angler_type_kayak_pontoon
    )

# dwg$interview <- 
#     CreelEstimateR::prep_dwg_interview_angler_types(
#       interview_fishing_time = dwg$interview,
#       study_design = params$study_design,
#       boat_type_collapse = params$boat_type_collapse,
#       fish_location_determines_type = params$fish_location_determines_type,
#       angler_type_kayak_pontoon = params$angler_type_kayak_pontoon
#     )
 
    

```

```{r manual data edit for vehicle counts}
#| include: false

 # add +1 "dummy" vehicle counts to trailer counts for 2022,2023

columns_to_keep <- setdiff(names(dwg$effort), c("count_type", "count_quantity"))

census <- dwg$effort |> 
  filter(tie_in_indicator == 1)
index_2021 <- dwg$effort |> 
  filter(tie_in_indicator == 0, year == 2021)
  
index_no_trailers <- dwg$effort |> 
  filter(year %in% c(2022,2023), tie_in_indicator == 0) |> 
   pivot_wider(
    id_cols = all_of(columns_to_keep),
    names_from = count_type,
    values_from = count_quantity
  ) |> 
  filter(
    `Trailers Only` == 0
  ) |> 
  pivot_longer(cols = c(`Vehicle Only`,`Trailers Only`), names_to = "count_type", values_to = "count_quantity")

# filter to index counts and trailer counts > 0
index_nonzero_trailers <- dwg$effort |>
  filter(year %in% c(2022,2023), tie_in_indicator == 0) |> 
  pivot_wider(
    id_cols = all_of(columns_to_keep),
    names_from = count_type,
    values_from = count_quantity
  ) |> 
    filter(
    `Trailers Only` > 0
  ) |> 
  mutate(
    `Vehicle Only` = `Vehicle Only`+ `Trailers Only`
  ) |> 
  pivot_longer(cols = c(`Vehicle Only`,`Trailers Only`), names_to = "count_type", values_to = "count_quantity")

dwg$effort <- index_nonzero_trailers |> 
  rbind(index_no_trailers, census, index_2021) |> 
  arrange(event_date, count_sequence)


```

### Angler interviews

```{r report_interview_totals}


report_tabs$interview_totals <- dwg$interview |> group_by(year) |> mutate(annual_total_interviews = n_distinct(interview_id)) |> 
  group_by(year, angler_final, annual_total_interviews) |> 
  summarise(
    total_angler_count = sum(angler_count),
    total_individual_count = sum(total_group_count),
    annual_interview_group_count = n_distinct(interview_id)
    )

# summarize the number of interviews obtained by trip_status (complete/incomplete trips)
report_tabs$trip_status_summary <- dwg$interview |> 
  count(year, angler_final, trip_status) |> 
  arrange(year, angler_final, trip_status) |> 
  group_by(year, angler_final) |> 
  mutate(
    total_interviews = sum(n),
    trip_status_proportion = round(n / total_interviews, 3) * 100
  )

# pull out the total number of interviews obatined in section 1
report_tabs$section_1_interviews <- dwg$interview |> filter(section_num == 1) |>  count(year, section_num) |> summarise(sum(n)) |> pluck() 

# pull out the total number of interviews obatined in section 3
report_tabs$section_3_interviews <- dwg$interview |> filter(section_num == 3) |>  count(year, section_num) |> summarise(sum(n)) |> pluck() 

# pull out the total number of interviews obatined in section 4
report_tabs$section_4_interviews <- dwg$interview |> filter(section_num == 4) |>  count(year, section_num) |> summarise(sum(n)) |> pluck() 

```

In `r report_tabs$interview_totals[1,1]`, sampling for interviews only occurred in section 2, where we obtained a total of `r report_tabs$interview_totals[1,3]` interviews, representing `r report_tabs$interview_totals[1,6]` `r report_tabs$interview_totals[1,2]` angler groups and `r report_tabs$interview_totals[2,6]` `r report_tabs$interview_totals[2,2]` angler groups. In `r report_tabs$interview_totals[3,1]`, a total of `r report_tabs$interview_totals[3,3]` interviews were obtained, representing `r report_tabs$interview_totals[3,6]` `r report_tabs$interview_totals[3,2]` angler group interviews and `r report_tabs$interview_totals[4,6]` `r report_tabs$interview_totals[4,2]` angler group interviews. In `r report_tabs$interview_totals[5,1]`, a total of `r report_tabs$interview_totals[5,3]` interviews were obtained, representing `r report_tabs$interview_totals[5,6]` `r report_tabs$interview_totals[5,2]` angler group interviews and `r report_tabs$interview_totals[6,6]` `r report_tabs$interview_totals[6,2]` angler group interviews. Across all years the vast majority of interviews were obtained in section 2. Relatively few interviews were obtained in the lower river represented by section 1 (*n* = `r report_tabs$section_1_interviews`) and the upper river, represented by section 3 (*n* = `r report_tabs$section_3_interviews`), and section 4 (*n* = `r report_tabs$section_4_interviews`). Sample sizes in these areas were likely limited by accessibility and logistical constraints for sampling staff. See Table \_ for a breakout of the number of interviews obtained by year, section, and angler type.

```{r}
#| fig.cap: "Figure _. The number of interviews obtained, categorized by bank (blue) and boat (orange) angling groups, per month during the Nisqually River salmon fishery from 2021 to 2023."


# number of angler groups interviewed (equivalent to # of interviews obtained)
 dwg$interview |> 
  count(year, month, angler_final, name = "interview_count") |> # count of interviews 
  rename(angler_type = angler_final) |>
  rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence()) |> 
  ggplot(aes(x = as.factor(Month), y = `Interview count`, fill = `Angler type`)) +
  geom_bar(stat = "identity", position = position_stack(), color = "black") +
  facet_wrap(~ Year, scales = "free_x") +
  # scale_x_discrete(limits = c('7', '8', '9', '10', '11', '12')) +
  scale_fill_manual(values = c("bank" = "steelblue", "boat" = "orange"), 
                    labels = c("Bank", "Boat")) +
  labs(
       x = "Month",
       y = "Number of angler group interviews",
       fill = "Angler type") +
  theme_bw()

```

```{r}
#| include: false

# summary stats describing interview sampling effort by year and month, with daily minimum, maximum, and mean number of interviews obtained
dwg$interview |>
  group_by(year, month, event_date) |>
  summarise(
    daily_count = n()
  ) |>
  group_by(year, month) |>
  summarise(
    mean = mean(daily_count),
    sd = sd(daily_count),
    min = min(daily_count),
    max = max(daily_count), .groups = "drop"
  ) |> 
  mutate(across(mean:sd, ~ round(., 0))) |> 
  gt(caption = "Summary stats describing angler sampling by year and month, with daily minimum, maximum, and mean number of interviews obtained.") 

```

```{r interview_and_catch}
#| eval: false
#| include: false

dwg$interview |>
  count(water_body, year, section_num, angler_final) |> 
  arrange(year, section_num, angler_final) |> 
  select(year, section_number = section_num, angler_type = angler_final, count = n) |> 
  rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence()) |> 
  gt(caption = "Table _. The number of interviews obtained during roving creel surveys, grouped by year, section, and angler type during the Nisqually River salmon fishery from 2021 to 2023.") 

```

```{r}
#| include: false
#| eval: false

# number of individuals within groups interviewed
 dwg$interview |>
  group_by(year, month, angler_final) |>
  summarise(
    total_angler_count = sum(angler_count)
    # ,
    # total_individual_count = sum(total_group_count) # this should be toggled by params$person_count_type! 
    ) |>
  rename(angler_type = angler_final) |>
  rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence()) |>
  ggplot(aes(x = as.factor(Month), y = `Total angler count`, fill = `Angler type`)) +
  geom_bar(stat = "identity", position = position_stack(), color = "black") +
  facet_wrap(~ Year, scales = "free_x") +
  scale_x_discrete(limits = c('7', '8', '9', '10', '11')) +
  scale_fill_manual(values = c("bank" = "steelblue", "boat" = "orange"),
                    labels = c("Bank", "Boat")) +
  labs(
       x = "Month",
       y = "Number of angler group members interviewed",
       fill = "Angler type") +
  theme_bw()

```

```{r, month_mean_trip_length}
#| include: false

interview_group_sum_count_year_month_angler <- dwg$interview |>
  filter(year %in% c(2022,2023)) |> 
  filter(previously_interviewed == 0|is.na(previously_interviewed)) |>
   mutate(
    total_group_count = if_else(is.na(total_group_count), angler_count, total_group_count)) |> 
  group_by(fishery_name, year, month, angler_final) |>
  summarize(
    n_individuals = sum(total_group_count)
  ) |>
  rename(angler_type = angler_final)


# sum of all angler hours reported in interviews
fishing_time_sum <- dwg$interview |> 
  select(-angler_type) |> 
  # filter(trip_status == "Complete", previously_interviewed == 0) |> # depends on whether field previsouly_interviewed was used!
  rename(angler_type = angler_final) |> 
  group_by(year, month, angler_type) |> 
  summarise(
    sum_fishing_hours = sum(fishing_time_total)
  )
  
# mean trip length for individuals, not total angler group 
# mean_trip_length_monthly <- dwg$interview |> 
#   select(-angler_type) |> 
#   # filter(trip_status == "Complete", previously_interviewed == 0) |> # depends on whether field previsouly_interviewed was used!
#  rename(angler_type = angler_final) |> 
#   filter(trip_status == "Complete") |> 
#   group_by(year, month, angler_type) |> 
#   summarize(
#     n_completed_angler_trips = n(),
#     mean_trip_length = mean(fishing_time),
#     sd = sd(fishing_time),
#     se = sd / sqrt(n_completed_angler_trips),
#     cv = sd / mean_trip_length
#   ) |> 
#   left_join(fishing_time_sum)

# mean trip length for individuals, not total angler group 
# mean_trip_length_monthly <- dwg$interview |> 
#   select(-angler_type) |> 
#   # filter(trip_status == "Complete", previously_interviewed == 0) |> # depends on whether field previsouly_interviewed was used!
#  rename(angler_type = angler_final) |> 
#   filter(trip_status == "Complete") |> 
#   group_by(year, month, angler_type) |> 
#   summarize(
#     n_completed_angler_trips = sum(angler_count),
#     mean_trip_length = mean(fishing_time),
#     sd = sd(fishing_time),
#     se = sd / sqrt(n_completed_angler_trips),
#     cv = sd / mean_trip_length
#   ) |> 
#   left_join(fishing_time_sum)


# summary statistics for angler group interviews
mean_group_trip_length_monthly <- dwg$interview |> 
  select(-angler_type) |> 
  filter(year %in% c(2022,2023)) |>
  # filter(trip_status == "Complete", previously_interviewed == 0) |> # depends on whether field previsouly_interviewed was used!
 rename(angler_type = angler_final) |> 
  filter(trip_status == "Complete") |> 
  group_by(year, month, angler_type) |> 
  summarize(
    n_completed_group_interviews = n(),
    mean_group_trip_length = mean(fishing_time_total),
    sd = sd(fishing_time_total),
    se = sd / sqrt(n_completed_group_interviews),
    cv = sd / mean_group_trip_length
  ) |>
  left_join(fishing_time_sum)


mean_individual_trip_length_monthly <- dwg$interview |> 
  select(-angler_type) |> 
  # filter(year %in% c(2022,2023)) |>
  # filter(trip_status == "Complete", previously_interviewed == 0) |> # depends on whether field previsouly_interviewed was used!
 rename(angler_type = angler_final) |> 
  filter(trip_status == "Complete") |> 
  group_by(year, month, angler_type) |> 
  summarize(
    n_completed_group_interviews = n(),
    mean_individual_trip_length = mean(fishing_time),
    # mean_group_trip_length = mean(fishing_time_total),
    sd = sd(fishing_time),
    # se = sd / sqrt(n_completed_group_interviews),
    # cv = sd / mean_individual_trip_length,
    sum_angler_count = sum(angler_count)
  ) |>
  left_join(fishing_time_sum)

mean_trip_length_annual <- dwg$interview |> 
  filter(trip_status == "Complete") |>
  group_by(angler_final) |> 
  summarize(
    n_completed_angler_trips = n(),
    # n_individuals = sum(total_group_count),
    mean_trip_length = mean(fishing_time),
    # , # the mean trip length for groups, which is the matching resolution at which interviews are obtained!
    sd = sd(fishing_time),
    # se = sd / sqrt(n_completed_angler_trips),
    cv = sd / mean_trip_length
  ) |> 
  rename(angler_type = angler_final) 


report_tabs$mean_individual_trip_length_monthly <- mean_individual_trip_length_monthly |> 
  mutate(
    mutate(across(mean_individual_trip_length:sum_fishing_hours, ~ round(., 2)))
  ) |> 
  rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence()) |>
  rename(SD = Sd)

```

```{r mean trip length plot}
#| include: true
#| fig.cap: "Figure _. Mean monthly trip time (± SD) of bank (blue) and boat (orange) angling groups surveyed on during the Nisqually River salmon fishery from 2021 to 2023."

# Generate bar plot with error bars
report_tabs$mean_individual_trip_length_monthly |> 
ggplot(aes(x = factor(Month), y = `Mean individual trip length`, fill = `Angler type`)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  geom_errorbar(aes(ymin = `Mean individual trip length` - SD, ymax = `Mean individual trip length` + SD), 
                position = position_dodge(0.9), width = 0.25) +
  labs(
    x = "Month",
    y = "Mean angler trip length (hours)",
    fill = "Angler type"
  ) +
  facet_wrap(~ Year, scales = "free_x") +
  # scale_x_discrete(limits = c('7', '8', '9', '10', '11', '12'))
  scale_fill_manual(values = c("bank" = "steelblue", "boat" = "orange"), 
                    labels = c("Bank", "Boat")) +
  theme_bw()

```

### Effort surveys

#### Index effort counts

```{r report_effort_index_totals}
#| include: false

# summarise the index count data
# not likely many useful insights from the individual counts
# monthly summary statistics to start 
report_tabs$effort_index_mean_daily <- dwg$effort |> 
  filter(tie_in_indicator == 0) |> 
  dplyr::group_by(year, month, section_num, event_date, count_sequence, count_type) |> #
  dplyr::summarise(count_index = sum(count_quantity), .groups = "drop") |> 
  dplyr::group_by(year, month, section_num, event_date, count_type) |> 
  dplyr::summarise(count_index_mean = mean(count_index), .groups = "drop")


report_tabs$effort_index_monthly_summary_stats <- report_tabs$effort_index_mean_daily |> 
  dplyr::group_by(year, month, count_type) |> 
  summarise(
    sample_size = n(),
    mean = mean(count_index_mean),
    standard_deviation = sd(count_index_mean),
    minimum = min(count_index_mean),
    maximum = max(count_index_mean)
  ) |> 
   mutate(across(where(is.numeric), round, digits = 2)) 

  
```

```{r effort index check}
#| eval: false
#| 
dwg$effort |> 
  filter(tie_in_indicator == 0) |> 
  dplyr::group_by(year, month, section_num, event_date, count_sequence, count_type) |> #
  dplyr::summarise(count_index = sum(count_quantity), .groups = "drop") |> 
  dplyr::group_by(year, month, section_num, event_date, count_type) |> 
  dplyr::summarise(count_index_mean = mean(count_index), .groups = "drop")

```

```{r effort_index_annual_stats}
#| include: false

report_tabs$effort_index_annual_summary_stats <- report_tabs$effort_index_mean_daily |> 
  dplyr::group_by(year, count_type) |> 
  summarise(
    sample_size = n(),
    mean = mean(count_index_mean),
    standard_deviation = sd(count_index_mean),
    minimum = min(count_index_mean),
    maximum = max(count_index_mean)
  ) |> 
   mutate(across(where(is.numeric), round, digits = 2)) 

# report_tabs$effort_index_annual_summary_stats <- report_tabs$effort_index_annual_summary_stats |> 
#   filter(year %in% c(2022,2023))
```

Annual patterns in counts of vehicles and trailers exhibited some variability among years during the monitoring period. In `r report_tabs$effort_index_annual_summary_stats[1,1]`, across all months the mean daily index effort count of boat trailers was `r report_tabs$effort_index_annual_summary_stats[1,4]` (SD = `r report_tabs$effort_index_annual_summary_stats[1,5]`, *n* = `r report_tabs$effort_index_annual_summary_stats[1,3]`) and the mean daily index effort count of vehicles was `r report_tabs$effort_index_annual_summary_stats[2,4]` (SD = `r report_tabs$effort_index_annual_summary_stats[2,5]`, *n* = `r report_tabs$effort_index_annual_summary_stats[2,3]`). In `r report_tabs$effort_index_annual_summary_stats[3,1]`, across all months the mean daily index effort count of boat trailers was `r report_tabs$effort_index_annual_summary_stats[3,4]` (SD = `r report_tabs$effort_index_annual_summary_stats[3,5]`, *n* = `r report_tabs$effort_index_annual_summary_stats[3,3]`) and the mean daily index effort count of vehicles was `r report_tabs$effort_index_annual_summary_stats[4,4]` (SD = `r report_tabs$effort_index_annual_summary_stats[4,5]`, *n* = `r report_tabs$effort_index_annual_summary_stats[4,3]`). In `r report_tabs$effort_index_annual_summary_stats[5,1]`, across all months the mean daily index effort count of boat trailers was `r report_tabs$effort_index_annual_summary_stats[5,4]` (SD = `r report_tabs$effort_index_annual_summary_stats[5,5]`, *n* = `r report_tabs$effort_index_annual_summary_stats[5,3]`) and the mean daily index effort count of vehicles was `r report_tabs$effort_index_annual_summary_stats[6,4]` (SD = `r report_tabs$effort_index_annual_summary_stats[6,5]`, *n* = `r report_tabs$effort_index_annual_summary_stats[6,3]`).

Monthly vehicle and trailer count data exhibited similar seasonal patterns in across the three years, with some differences in variability related to time period and count type. Mean daily index effort counts in August and September of 2021 were higher than in corresponding months in 2022 and 2023. Importantly, we did not conduct census effort counts in 2021, which limits inferences that could be made regarding true difference in angling effort across years. In each month, the standard deviation of mean daily vehicle counts was larger than that observed for trailer counts, indicating more variability in index effort counts of vehicles than boat trailers.

```{r effort_index_annual_stats_table}
#| include: false
#| eval: false


report_tabs$effort_index_annual_summary_stats |> 
  # gt()|> 
  knitr::kable(caption = "Annual summary statistics describing index effort counts, where survey staff conducted standardized counts of vehicles and boat trailers to estimate angler abundance in the fishery area. Statistics include the sample size of mean daily counts, standard deviation, and minimum and maximum mean daily counts.")
```

```{r effort_index_box_whisker_plot}
#| fig.cap: "Figure _. Box and whisker plot of mean daily index effort counts, categorized by vehicle (green) and boat trailer (red) count types, per month on the Nisqually River from 2021 to 2023. The lower and upper hinges correspond to the first and third quartiles (the 25th and 75th percentiles). The upper whisker extends from the hinge to the largest value no further than 1.5 * IQR from the hinge (where IQR is the inter-quartile range, or distance between the first and third quartiles). The lower whisker extends from the hinge to the smallest value at most 1.5 * IQR of the hinge. Data beyond the end of the whiskers are plotted individually."
#| fig-width: 8

# Generate boxplot
report_tabs$effort_index_mean_daily |> 
rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence()) |> 
ggplot(aes(x = factor(Month), y = `Count index mean`, fill = `Count type`)) +
  geom_boxplot(color = "black") +
  labs(x = "Month",
       y = "Effort index count (mean daily)") +
  facet_wrap(~ Year, scales = "free_x") +
  theme_bw()
```

```{r effort_index_line_plot}
#| fig.cap: "Figure _. Mean daily index effort counts of vehicles and trailers from 2021 through 2023 in the Nisqually River. July counts are not shown for 2022 and 2023 for consist"
#| fig-width: 8

# Generate boxplot
report_tabs$effort_index_mean_daily |> 
filter(month %in% c(7,8,9,10,11)) |> 
rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence()) |> 
ggplot(aes(x = `Event date`, y = `Count index mean`, color = `Count type`)) +
  geom_point() +
  labs(x = "Date",
       y = "Effort index count (mean daily)") +
  facet_wrap(~ Year, scales = "fixed") +
  theme_bw()
```

#### Census effort counts

```{r fix_count_type_name}
#| include: false

dwg$effort <- dwg$effort |> 
  mutate(count_type = str_replace(count_type, "Shore", "Bank"))

```

```{r report_effort_census_totals}

report_tabs$effort_census_events  <- dwg$effort |> 
  filter(tie_in_indicator == 1) |>
  select(year, event_date, count_type, count_quantity) |> 
  group_by(year, event_date, count_type) |>  # summarize over two sections 
  summarise(count_quantity = sum(count_quantity), .groups = "drop") |> 
  pivot_wider(names_from = count_type, values_from = count_quantity) |> 
  rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence())


report_tabs$effort_census_annual_summary <- report_tabs$effort_census_events |> 
  count(Year) #add to this as needed

```

```{r report_effort_census_percentage_angler_type}
#| include: false

census_effort_proportions <- dwg$effort |>
  filter(tie_in_indicator == 1) |> 
  group_by(year) |> 
  mutate(total = sum(count_quantity)) |> 
  ungroup() |> 
  group_by(year, count_type, total) |>
  summarise(
    count_sum = sum(count_quantity),
    percentage = round(count_sum/total, 2),
    .groups = "drop"
  ) |> 
  distinct() |> 
  select(-c(total)) |> 
  arrange(year, desc(percentage))

census_effort_proportions |> 
  rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence()) |> 
  gt()

```

In `r report_tabs$effort_census_annual_summary[1,1]` we conducted `r report_tabs$effort_census_annual_summary[1,2]` census effort counts of anglers. In `r report_tabs$effort_census_annual_summary[2,1]` we conducted `r report_tabs$effort_census_annual_summary[2,2]` census effort counts of anglers. Proportions of observed angler types were similar across years. Cumulative census counts of two angler types, 1) bank anglers without a vessel, and 2) boat anglers using a motorized boat, drift boat, or raft made up `r census_effort_proportions[1,4]*100`% and `r (census_effort_proportions[2,4]+census_effort_proportions[3,4])*100`% of observed anglers in 2022, respectively, and `r census_effort_proportions[8,4]*100`% and `r (census_effort_proportions[9,4]+census_effort_proportions[10,4])*100`% of observed anglers in 2023 (Figure ).

```{r census_count_histogram}

# bar_plot <- ggplot(dwg$effort |>
#          filter(tie_in_indicator == 1), aes(x = factor(year), y = count_quantity, fill = count_type)) +
#   geom_bar(stat = "identity") +
#   labs(
#        x = "Year",
#        y = "Cumulative angler census count",
#        fill = "Angler type") +
#   theme_bw()


```

```{r census_count_totals_barplot}
#| fig.cap: "Figure _. Stacked barplots with cumulative counts of anglers observed during census effort count surveys in the Nisqually River in 2022 and 2023. The table contains cumulative counts, with each row grouped by angler type and columns grouping data by year."

# Prepare summary table
df_summary <- dwg$effort |> 
  filter(tie_in_indicator == 1) |>  
  group_by(year, count_type) |> 
  summarise(count_sum = sum(count_quantity), .groups = "drop") |> 
  pivot_wider(names_from = year, values_from = count_sum, values_fill = 0) |> 
  arrange(desc(count_type))  # Ensure consistency in order

table_plot <- ggplot(df_summary |> 
                        pivot_longer(cols = -count_type, 
                                     names_to = "Year", values_to = "Total_Count"), 
                     aes(x = Year, y = count_type, label = Total_Count)) +
  geom_tile(alpha = 0.1) +
  geom_text(size = 3) +
  theme_void() +
  # labs(title = "Total Counts by Year and Angler Type") +
  theme(
    # plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(hjust = 0.5),
    axis.text.y = element_text(size = 9)
  )


# Create the stacked bar plot
bar_plot <- ggplot(dwg$effort |>  
                     filter(tie_in_indicator == 1), 
                   aes(x = factor(year), y = count_quantity, fill = count_type)) +
  geom_bar(stat = "identity") +
  labs(
       x = "Year",
       y = "Census angler count (cumulative)",
       fill = "Angler type") +
  theme_bw() +
  theme(
    panel.grid.major = element_line(color = "gray80", size = 0.5),  # Major grid lines
    panel.grid.minor = element_line(color = "gray90", size = 0.3),  # Minor grid lines
    panel.background = element_rect(fill = "white")  # Keeps background clean
  )

grid.arrange(bar_plot, table_plot, heights = c(8,4))
```

```{r effort_estimates}
#| include: false

# bss effort by fishery/period/angler type
bss_effort_mean <- creel_estimates$stratum |>
  filter(
    model_type == "BSS",
    estimate_category == "effort",
    estimate_type == "mean"
    )


avg_effort <- bss_effort_mean |> 
  group_by(year, month, angler_type_name) |> #sum across days in a week
  summarise(mean_total_effort = round(sum(estimate_value), 1)) |>
  rename(angler_type = angler_type_name) |> 
  mutate(
    angler_type = str_to_lower(word(angler_type))
  )


```

```{r plot_mean_effort}
#| eval: false
#| include: false

plot_effort_avg_effort <- avg_effort |> 
  ggplot(aes(x = month, y = mean_total_effort, fill = angler_type)) +
  geom_col() +
  facet_wrap(~year) +
  theme_bw() +
  scale_x_continuous(breaks = unique(avg_effort$month)) +
  labs(x = "Week", y = "Mean Effort (hours)")

plot_effort_avg_effort

```

### Angler trips and sampling rate

#### Estimated number of angler trips

Across all months and years, the mean trip length of banks anglers was `r round(mean_trip_length_annual[1,3], 2)` hours (SD = `r round(mean_trip_length_annual[1,4], 2)`, CV = `r round(mean_trip_length_annual[1,5], 2)`), and the mean trip length of boat anglers was `r round(mean_trip_length_annual[2,3], 2)` hours (SD = `r round(mean_trip_length_annual[2,4], 2)`, CV = `r round(mean_trip_length_annual[2,5], 2)`). Bank anglers on average fished for less time than boat anglers and exhibited more variability in trip length time. Within each angler type, estimates of mean trip length exhibited little variability across months and lacked discernible temporal trends (Figure 9).

```{r angler_trip_tables}
#| include: false
#| eval: true

# calculate estimated number of angler trips - total effort / mean trip length (for groups, not individuals)
# to do! create one table with information for group and individual anglers (fishing time total / fishing time)

angler_trips <- mean_individual_trip_length_monthly |> 
  ungroup() |> 
  left_join(avg_effort, by = c("year", "month", "angler_type")) |> 
  mutate(
    estimated_angler_trips = mean_total_effort / mean_individual_trip_length
  ) |> 
  left_join(interview_group_sum_count_year_month_angler, by = c("year", "month", "angler_type")) |> 
  mutate(
    angler_type = str_to_sentence(angler_type)
   ) |> 
  mutate(
    sampled_angler_trip_proportion =  sum_angler_count / estimated_angler_trips,
    sampled_angler_hours_proportion = sum_fishing_hours / mean_total_effort
  )


angler_trips_angler_type_aggregate <- mean_individual_trip_length_monthly |> 
  ungroup() |> 
  left_join(avg_effort, by = c("year", "month", "angler_type")) |> 
  mutate(
    estimated_angler_trips = mean_total_effort / mean_individual_trip_length
  ) |> 
  left_join(interview_group_sum_count_year_month_angler, by = c("year", "month", "angler_type")) |> 
  mutate(
    angler_type = str_to_sentence(angler_type)
   ) |> 
  # ungroup() |> 
  group_by(year, month, angler_type) |> 
  summarise(
    total_individuals = sum(sum_angler_count), 
    total_trips = sum(estimated_angler_trips),
    sampled_proportion =  total_individuals / total_trips, .groups = "drop"
  )



angler_trips_aggregate <- mean_individual_trip_length_monthly |> 
  filter(!year == 2021) |> 
  ungroup() |> 
  left_join(avg_effort, by = c("year", "month", "angler_type")) |> 
  mutate(
    estimated_angler_trips = mean_total_effort / mean_individual_trip_length
  ) |> 
  left_join(interview_group_sum_count_year_month_angler, by = c("year", "month", "angler_type")) |> 
  mutate(
    angler_type = str_to_sentence(angler_type)
   ) |> 
  # ungroup() |> 
  group_by(year, angler_type) |> 
  summarise(
    total_individuals = sum(sum_angler_count), 
    total_trips = sum(estimated_angler_trips),
    sampled_proportion =  total_individuals / total_trips, .groups = "drop"
  )


angler_trips_aggregate_final <- mean_individual_trip_length_monthly |> 
  ungroup() |> 
  left_join(avg_effort, by = c("year", "month", "angler_type")) |> 
  mutate(
    estimated_angler_trips = mean_total_effort / mean_individual_trip_length
  ) |> 
  left_join(interview_group_sum_count_year_month_angler, by = c("year", "month", "angler_type")) |> 
  mutate(
    angler_type = str_to_sentence(angler_type)
   ) |> 
  # ungroup() |> 
  group_by(angler_type, year) |> 
  summarise(
    total_individuals = sum(sum_angler_count), 
    total_trips = sum(estimated_angler_trips),
    total_interview_hours = sum(sum_fishing_hours),
    total_estimated_effort = sum(mean_total_effort),
    sampled_proportion_trips =  total_individuals / total_trips,
    sampled_proportion_hours = total_interview_hours / total_estimated_effort,
    .groups = "drop"
  )

# |>
  # gt()

```

#### Angler Sampling rate

Using the monthly mean trip time and estimated seasonal total angling effort (angler-hours), we estimated the number of bank and boat angler trips. In 2022, an estimated `r 16202 + 1109` angling trips occurred *(n*~bank~ = `r round(angler_trips_aggregate[1,4], 0)`, *n*~boat~ = `r round(angler_trips_aggregate[2,4], 0)`. In 2023, an estimated `r 17149 + 1227` angling trips occurred (*n*~bank~ = `r round(angler_trips_aggregate[3,4], 0)`, *n*~boat~ = `r round(angler_trips_aggregate[4,4], 0)`. Given relative consistency in monthly mean trip length estimates (Figure 9), the estimated number of angler-trips per month (Figure 10) followed a similar pattern to the estimated seasonal total angler effort.

```{r angler_trip_table}
#| eval: false
#| include: false

angler_trips |>
  drop_na(mean_total_effort) |> 
  mutate(across(mean_individual_trip_length:sd, ~ round(., 2))) |>
  mutate(across(sampled_angler_trip_proportion:sampled_angler_hours_proportion, ~ round(., 2))) |> 
  gt()
  # knitr::kable(caption = "Table _. Table with angler trip information.")

```

{{< pagebreak >}}

```{r angler_trip_barplots}
#| fig.cap: "Figure _. The estimated number of individual angler trips, categorized by bank (blue) and boat (orange) angler types, per month on the Humptulips River from September 1, 2022 to November 29, 2023. The fishery was closed through emergency regulations from October 8, 2022 to October 28, 2022 due to drought conditions. The fishery was not open in December 2023."

# number of angler groups interviewed (equivalent to # of interviews obtained)
angler_trips |> 
    drop_na(mean_total_effort) |> 
  # count(year, month, angler_final, name = "interview_count") |> # count of interviews 
  # rename(angler_type = angler_final) |>
  # rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence()) |> 
  ggplot(aes(x = as.factor(month), y = estimated_angler_trips, fill = angler_type)) +
  geom_bar(stat = "identity", position = position_stack(), color = "black") +
  facet_wrap(~ year, scales = "free_x") +
  # scale_x_discrete(limits = c('7', '8', '9', '10', '11', '12'))
  scale_fill_manual(values = c("Bank" = "steelblue", "Boat" = "orange"), 
                    labels = c("Bank", "Boat")) +
  labs(
       x = "Month",
       y = "Estimated number of individual angler trips",
       fill = "Angler type") +
  theme_bw()

```

```{r linear_fit}
#| include: false
# fit simple linear regression models: estimated # angler trips (monthly) ~ # of interviews (monthly) 

nested_models <- angler_trips |> 
  drop_na(mean_total_effort) |> 
  nest(data = -angler_type) |> 
  mutate(
    model = map(data, ~ {
      lm_spec <- linear_reg() |> 
        set_engine("lm")

      workflow() |> 
        add_model(lm_spec) |> 
        add_formula(estimated_angler_trips ~ n_individuals) |> 
        fit(data = .x)
    }),
    predictions = map2(model, data, ~ {
      preds <- predict(.x, new_data = .y)
      bind_cols(.y, pred = preds$.pred)
    }),
    model_summary = map(model, ~ {
      fitted_model <- extract_fit_parsnip(.x) # Extract fitted model
      glance(fitted_model)
    })
  )
```

```{r extract_coefficients}
#| include: false

# Extract coefficients and create parameterized model statements using format in Bentley et al. 2015
# "the estimated number of anglers per month was parameterized as y(Number of Bank Anglers) = 6.24(Bank Angler Interviews) and y(Number of Boat Anglers) = 2.71(Boat Angler Interviews)."

model_statements <- map_dfr(nested_models$model, ~ {
  # Extract the underlying fitted parsnip model
  fit <- extract_fit_parsnip(.x)
  # Tidy extraction of coefficients
  coefs <- tidy(fit)
  
  # Retrieve the intercept and slope
  intercept <- coefs %>% filter(term == "(Intercept)") %>% pull(estimate)
  slope <- coefs %>% filter(term == "n_individuals") %>% pull(estimate)
  
  # Retrieve the group identifier from the data
  current_group <- unique(.x$data$angler_type) # this isn't working to extract angler_type
  
  tibble(
    intercept = round(intercept, 2),
    slope = round(slope, 2),
    angler_type = current_group,
    statement = paste0(
      "The linear model fit for ", current_group, " anglers is parameterized as: ",
      "y(Estimated Angler Trips) = ", round(slope, 2), " (Angler Interviews) + ", round(intercept, 2)
    )
  )
})

# Print the parameterized model statements

```

```{r}
# The estimated number of monthly individual angler trips was related to the observed number of individual anglers that were interviewed during creel surveys (Figure ; Bank: p < `r round(nested_models$model_summary[[1]]$p.value, 6)`, R^2^ = `r round(nested_models$model_summary[[1]]$r.squared, 2)`; Boat: p < `r round(nested_models$model_summary[[2]]$p.value, 4)`, R^2^ = `r round(nested_models$model_summary[[2]]$r.squared, 2)`). The linear model fit for bank anglers was parameterized as: y(Estimated Angler Trips) = `r model_statements$slope[1]` (Angler Interviews) `r model_statements$intercept[1]`. The linear model fit for boat anglers was parameterized as: y(Estimated Angler Trips) = `r model_statements$slope[2]` (Angler Interviews) `r model_statements$intercept[2]`. Therefore, approximately `r round((1/model_statements$slope[1])*100, 0)`% of bank anglers were interviewed, and approximately `r round((1/model_statements$slope[2])*100, 0)`% of boat anglers were interviewed, across all months and years.
```

```{r model_fit_and_observed_data_plot}
#| eval: false
#| fig.cap: "Figure _. The estimated number of individual-level angler trips as a function of the total number of individuals within anglers groups interviewed per month, categorized by bank (blue) and boat (orange) angling groups, on the Humptulips River from September 1, 2022 to November 29, 2023. The fishery was closed through emergency regulations from October 8, 2022 to October 28, 2022 due to drought conditions. The fishery was not open in December 2023."


# Unnest the predictions so that all groups are in one dataframe for plotting:
predicted_data <- nested_models |> 
  unnest(predictions)

max_y <- max(predicted_data$estimated_angler_trips, predicted_data$pred, na.rm = TRUE)

# Visualize: scatterplot of observed data with the fitted regression lines
# trips_by_interviews_plot <- 
  
ggplot(predicted_data, aes(x = n_individuals, y = estimated_angler_trips, color = angler_type)) +
  geom_point(size = 2, alpha = 0.8) +        # Observed data
  geom_line(aes(y = pred), size = 1) +         # Fitted regression line
  # facet_wrap(~ angler_type) +                  # Separate plots for each group
  labs(
    color = "Angler type",
    x = "Number of individuals in interviewed groups (monthly)",
    y = "Estimated number of individual-level angler trips (monthly)"
  ) +
  # facet_wrap(.~angler_type) +
  # scale_fill_manual(values = c("Bank" = "steelblue", "Boat" = "orange"), labels = c("Bank", "Boat")) +
  scale_color_manual(values = c("Bank" = "steelblue", "Boat" = "orange"), labels = c("Bank", "Boat")) +
  # scale_y_continuous(limits = c(0, 2000)) +
  # scale_x_continuous(limits = c(0,440)) +
  theme_bw() 

# trips_by_interviews_plot
# predicted_data |> filter(angler_type == "Bank") |> 
# ggplot(aes(x = n_individuals, y = estimated_angler_trips, color = angler_type)) +
#   geom_point(size = 2, alpha = 0.8) +        # Observed data
#   geom_line(aes(y = pred), size = 1) +         # Fitted regression line
#   # facet_wrap(~ angler_type) +                  # Separate plots for each group
#   labs(
#     color = "Angler type",
#     x = "Number of individuals in interviewed groups (monthly)",
#     y = "Estimated number of individual-level angler trips (monthly)"
#   ) +
#   # facet_wrap(.~angler_type) +
#   # scale_fill_manual(values = c("Bank" = "steelblue", "Boat" = "orange"), labels = c("Bank", "Boat")) +
#   scale_color_manual(values = c("Bank" = "steelblue", "Boat" = "orange"), labels = c("Bank", "Boat")) +
#   # scale_y_continuous(limits = c(0, 2000)) +
#   # scale_x_continuous(limits = c(0,440)) +
#   theme_bw() 

# predicted_data |> filter(angler_type == "Boat") |>
# ggplot(aes(x = n_individuals, y = estimated_angler_trips, color = angler_type)) +
#   geom_point(size = 2, alpha = 0.8) +        # Observed data
#   geom_line(aes(y = pred), size = 1) +         # Fitted regression line
#   # facet_wrap(~ angler_type) +                  # Separate plots for each group
#   labs(
#     color = "Angler type",
#     x = "Number of individuals in interviewed groups (monthly)",
#     y = "Estimated number of individual-level angler trips (monthly)"
#   ) +
#   # facet_wrap(.~angler_type) +
#   # scale_fill_manual(values = c("Bank" = "steelblue", "Boat" = "orange"), labels = c("Bank", "Boat")) +
#   scale_color_manual(values = c("Bank" = "steelblue", "Boat" = "orange"), labels = c("Bank", "Boat")) +
#   # scale_y_continuous(limits = c(0, 2000)) +
#   # scale_x_continuous(limits = c(0,440)) +
#   theme_bw() 
```

```{r}
# Unnest predictions
predicted_data <- nested_models |>  
  unnest(predictions)

# Determine maximum y-value for consistency across plots
max_y <- max(predicted_data$estimated_angler_trips, predicted_data$pred, na.rm = TRUE)

# Generate separate plots for each angler type

ggplot(predicted_data, aes(x = n_individuals, y = estimated_angler_trips, color = angler_type)) +
  geom_point(size = 2, alpha = 0.8) +  # Observed data
  geom_line(aes(y = pred), size = 1) +  # Fitted regression line
  geom_text(aes(label = month), vjust = -0.5, size = 3.5, color = "black", check_overlap = TRUE) +  # Add month labels
  facet_wrap(~ angler_type + year, scales = "free") +  # Allow scales to vary by angler type
  labs(
    color = "Angler type",
    x = "Number of individuals in interviewed groups (monthly)",
    y = "Estimated number of individual-level angler trips (monthly)"
  ) +
  scale_color_manual(values = c("Bank" = "steelblue", "Boat" = "orange"),
                     labels = c("Bank", "Boat")) +
  theme_bw()
```

```{r sample_rate_vs_estimated_effort_plot}
#| eval: true
#| fig.cap: "Figure _. The estimated sampling rate of anglers, calculated as a ratio of angler hours directly sampled in interviews to estimated total angling effort, as a function of the total number of individuals within anglers groups interviewed per month, categorized by bank (blue) and boat (orange) angling groups."
# sample_rate_effort_plot  <- 
  


# The proportion of all individual anglers (i.e., angler trips for bank and boat anglers combined) that were interviewed by a sampler varied among months between `r round(min(angler_trips_aggregate$sampled_proportion),2)` in September 2022 and `r round(max(angler_trips_aggregate$sampled_proportion),2)` in December 2022. Given that our sampling in terms of survey days of staffing was relatively stable across years and months (Figure ), variation in our sampling rate of anglers was likely related to seasonal dynamics in angler effort (Figure ). 


ggplot(angler_trips, aes(x = mean_total_effort, y = sampled_angler_trip_proportion, color = angler_type)) +
  geom_point(size = 2, alpha = 0.8) +        # Observed data
  labs(
    # size = 4,
    color = "Angler type",
    x = "Estimated monthly angler effort (angler-hours)",
    y = "Estimated sample rate of anglers"
  ) +
  scale_color_manual(values = c("Bank" = "steelblue", "Boat" = "orange"), labels = c("Bank", "Boat")) +
   scale_y_continuous(
   limits = c(0, 0.45)
   # ,                         # Set the y-axis limits
   # breaks = pretty_breaks(n = 10)                # Create approximately 10 nice breaks
  ) +
  theme_bw() 

# sample_rate_effort_plot
# grid.arrange(trips_by_interviews_plot, sample_rate_effort_plot
#              , heights = c(8,8)
#              )


# angler_trips |> 
# ggplot(aes(mean_total_effort, sampled_proportion)) +
#   geom_point()

```
