---
title: "Creel survey sampling evaluation"
execute:
    echo: false
author: "Evan Booher"
format:
  docx: 
    reference-doc: "WDFW_Report_Template.docx"
    fig-dpi: 600
editor: visual
params:
  focal_waterbody: "Nisqually River"
  study_design: "Standard"  
  boat_type_collapse: "Yes"
  fish_location_determines_type: "No"
  angler_type_kayak_pontoon: "bank" 
  person_count_type: "angler"
  period_pe: "week"
  period_bss: "day"
  days_wkend: "!r c('Saturday', 'Sunday')"
  day_length_expansion: "night closure"
  min_fishing_time: 0.5
---

```{r, echo = FALSE}
#| echo: false

library(CreelEstimateR)
library(creelutils)
library(tidyverse)
library(here)
library(gt)
library(kableExtra)
library(flextable)
```

```{r}
#| echo: false
#| include: false

#lists that contain creel datasets downloaded from data.wa.gov
dwg <- creelutils::get_fishery_data(fishery_names = "Nisqually salmon", year = 2021:2023)

# add date grouping variables
dwg$interview <- dwg$interview |> 
  mutate(
    month = lubridate::month(.data$event_date),
    year = lubridate::year(.data$event_date),
    week = lubridate::week(.data$event_date)
  )

# add date grouping variables
dwg$effort <- dwg$effort |>
  mutate(
    month = lubridate::month(.data$event_date),
    year = lubridate::year(.data$event_date),
    week = lubridate::week(.data$event_date)
  )

```

```{r report_tables}
#| echo: false

# empty list of objects for report 
report_tabs <- list()

```

```{r read in all fishery dates}
#| echo: false

# survey frame representing all days open and closed in each fishery year created these objects in separate instances of CreelEstimates code using prep_days() and export them to data folder HumptulipsCreel repo
# Specify file paths
file_paths <- c("cleaned_data/days_Nisqually_2021.Rds",
                "cleaned_data/days_Nisqually_2022.Rds",
                "cleaned_data/days_Nisqually_2023.Rds")

report_tabs$all_dates <- file_paths |> 
  map(read_rds) |> 
  bind_rows()

```

```{r survey day evaluation assuming section 1 represents entire fishery}
#| echo: false

# data wrangling to evaluate survey coverage - version aggregating over sections. If closure structure differed by section, do not use this code chunk, see chunk below including all sections

# use the effort data as the basis to determine if a creel survey occurred. This is imperfect and could be improved with a more exhaustive "check" for criteria across event, effort, and interview data components to determine if a creel survey was completed
report_tabs$survey_dates <- dwg$effort |> 
  filter(tie_in_indicator == 0) |>  
  distinct(event_date) |> 
  mutate(
    surveyed = TRUE
  )

# perform an expanding left join of the full survey frame represented by the all_dates object to the survey_dates object, pivot long the open and surveyed status columns with True/False values 
report_tabs$combined_dates <- report_tabs$all_dates |> 
  left_join(report_tabs$survey_dates, by = "event_date") |> 
   mutate(across(
    .cols = contains("surveyed"),
    .fns = ~ replace_na(.x, FALSE)
  )) |> 
  select(-c(open_section_2)) |> 
  rename(open_section = open_section_1) |> 
  select(event_date, starts_with(c("open", "surveyed"))) |> 
  pivot_longer(cols = starts_with(c("open", "surveyed")), 
               names_to = c(".value")) |> 
  mutate(year = year(event_date), month = month(event_date)) 

# Across the entire monitoring period, evaluate the number of open fishing days, survey days, proportion of open fishing days surveyed, and the number of closure days
report_tabs$survey_dates_summary <- report_tabs$combined_dates |> 
  summarise(
    survey_days = sum(across(starts_with("surveyed"), ~ .x), na.rm = TRUE),
    closure_days = sum(across(starts_with("open"), ~ !.x), na.rm = TRUE),
    total_fishing_days = n() - closure_days,
    proportion_surveyed = round(survey_days/total_fishing_days, 2)
  )

# Grouping for year and month, evaluate the number of open fishing days, survey days, proportion of open fishing days surveyed, and the number of closure days
report_tabs$combined_dates_monthly_p <- report_tabs$combined_dates |> 
  select(event_date, open_section) |> # assumes by default single section section_num = 1
  left_join(report_tabs$survey_dates, by = "event_date") |> 
   mutate(across(
    .cols = contains("surveyed"),
    .fns = ~ replace_na(.x, FALSE)
  )) |> 
  select(event_date, starts_with(c("open", "surveyed"))) |> 
  mutate(year = year(event_date), month = month(event_date)) |> 
  group_by(year, month) |>
summarise(
    survey_days = sum(across(starts_with("surveyed"), ~ .x), na.rm = TRUE),
    closure_days = sum(across(starts_with("open"), ~ !.x), na.rm = TRUE),
    total_fishing_days = n() - closure_days,
    proportion_surveyed = round(survey_days/total_fishing_days, 2),
    greater_than_50 = if_else(proportion_surveyed >= 0.5, 1, 0),
    less_than_50 = if_else(proportion_surveyed < 0.2, 1, 0)
  )


```

```{r survey day evaluation w/ sections}
#| echo: false
#| eval: false

# version accounting for sections - current version only set up for fisheries w/ 2 sections named 1,2 - not dynamic! 
# data wrangling to evaluate the number of open fishing days, survey days, proportion of open fishing days surveyed, and the number of closure days per month - version aggregating over sections. Use this version if closure structure differed by section

# use the effort data as the basis to determine if a creel survey occurred. This is imperfect and could be improved with a more exhaustive "check" for criteria across event, effort, and interview data components to determine if a creel survey was completed
report_tabs$survey_dates <- dwg$effort |> 
  filter(tie_in_indicator == 0) |>  
  distinct(event_date, section_num) |> 
  mutate(
    surveyed = TRUE
  ) |> 
  pivot_wider(names_from = section_num, values_from = surveyed, values_fill = FALSE) |> 
  select(event_date, surveyed_section_1 = 2, surveyed_section_2 = 3) # not dynamic for other fisheries!!

# perform an expanding left join of the full survey frame represented by the all_dates object to the survey_dates object, pivot long the open and surveyed status columns with True/False values 
report_tabs$combined_dates <- report_tabs$all_dates |> 
  left_join(report_tabs$survey_dates, by = "event_date") |> 
   mutate(across(
    .cols = contains("surveyed"),
    .fns = ~ replace_na(.x, FALSE)
  )) |> 
  select(event_date, starts_with(c("open", "surveyed"))) |> 
  pivot_longer(cols = starts_with(c("open", "surveyed")), 
               names_to = c(".value", "section_num"), 
               names_pattern = "(.*)_section_(.*)") |> 
  mutate(year = year(event_date), month = month(event_date)) 

# Across the entire monitoring period, evaluate the number of open fishing days, survey days, proportion of open fishing days surveyed, and the number of closure days
report_tabs$combined_dates |> 
  summarise(
    survey_days = sum(across(starts_with("surveyed"), ~ .x), na.rm = TRUE),
    closure_days = sum(across(starts_with("open"), ~ !.x), na.rm = TRUE),
    total_days = n(),
    proportion_surveyed = round(survey_days/total_days, 2),
    proportion_closed = round(closure_days/total_days, 2)
  )

# Grouping for year, month, section_num, evaluate the number of open fishing days, survey days, proportion of open fishing days surveyed, and the number of closure days
report_tabs$survey_dates_monthly_p <- report_tabs$combined_dates |> 
  group_by(year, month, section_num) |>
  summarise(
    survey_days = sum(across(starts_with("surveyed"), ~ .x), na.rm = TRUE),
    closure_days = sum(across(starts_with("open"), ~ !.x), na.rm = TRUE),
    total_days = n(),
    proportion_surveyed = round(survey_days/total_days, 2),
    proportion_closed = round(closure_days/total_days, 2)
  ) |> 
   rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence())

```

# Results

## Creel survey sampling evaluation

### Survey days and closures

Over the study period, a roving creel survey was conducted on the `r params$focal_waterbody` on `r report_tabs$survey_dates_summary[1]` out of the `r report_tabs$survey_dates_summary[3]` (`r round((report_tabs$survey_dates_summary[1]/report_tabs$survey_dates_summary[3]*100), 0)`%) days that the fishery was open. At the monthly time step the sampled proportion of open fishing days among months was relatively consistent, as `r sum(report_tabs$combined_dates_monthly_p$greater_than_50)` of the `r nrow(report_tabs$combined_dates_monthly_p)` months had between 50% and 60% of days sampled, and one month, November 2023, where `r round(report_tabs$combined_dates_monthly_p[7,6]*100, 0)` of open fishing were sampled (Figure \_). In October 2022 the number of open fishing days were reduced due to an emergency closure from October 8, 2022 to October 28, 2022, due to low flow conditions. In December 2022 the upper section of the river from the Highway 101 Bridge upstream to the confluence of the East Fork and West Fork Humptulips Rivers was closed from December 1, 2022 to December 31st, and the lower section of the river from the mouth to the Highway 101 bridge was closed from December 17, 2022 to December 31, 2022. In December 2023 the entire fishery area was closed by emergency rule.

```{r}
#| echo: false

report_tabs$combined_dates_monthly_p |> 
  select(-c(greater_than_50, less_than_50)) |> 
  select(month, open_fishing_days = total_fishing_days, survey_days, proportion_surveyed, closure_days) |> 
   rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence()) |> 
  gt(caption = "Table _. The number of open fishing days, survey days, proportion of open fishing days surveyed, and the number of closure days per month during the 2022 and 2023 salmon fishery in the Humptulips River. Surveys were conducted from September 1, 2022 to November 29, 2023. The fishery was closed through emergency regulations from October 8-28, 2022 due to drought conditions. The upper section of the river from the Highway 101 Bridge upstream to the confluence of the East Fork and West Fork Humptulips Rivers was closed from December 1-31, 2022, and the lower section of the river from the mouth to the Highway 101 bridge  was closed from December 17-31, 2022. In December 2023 the fishery was closed by emergency rule.")
```

```{r}
#| echo: false
#| fig.cap: "Figure _. The number of open fishing days, survey days, and closure days per month during the 2022 and 2023 Humptulips River salmon fisheries. Surveys were conducted from September 1, 2022 to November 29, 2023. The fishery was closed through emergency regulations from October 8-28, 2022 due to drought conditions. The upper section of the river from the Highway 101 Bridge upstream to the confluence of the East Fork and West Fork Humptulips Rivers was closed from December 1-31, 2022, and the lower section of the river from the mouth to the Highway 101 bridge  was closed from December 17-31, 2022. In December 2023 the fishery was closed by emergency rule."


report_tabs$combined_dates_monthly_p |>
  select(year, month, open_fishing_days = total_fishing_days, closure_days, survey_days) |> 
  pivot_longer(cols = c(open_fishing_days, survey_days, closure_days),
               names_to = "day_type",
               values_to = "days") |> 
  mutate(day_type = factor(day_type, levels = c("open_fishing_days", "survey_days", "closure_days", "total_days"))) |> 
  ggplot(aes(x = as.factor(month), y = days, fill = day_type)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  facet_wrap(~ year, scales = "free_x") +
  labs(x = "Month",
       y = "Number of days",
       fill = "Day category") +
  scale_fill_manual(values = c("open_fishing_days" = "steelblue", "survey_days" = "orange", "closure_days" = "red"),
                    labels = c("open_fishing_days" = "Open fishing days", "survey_days" = "Survey days", "closure_days" = "Closure days")
                    ) +
  theme_bw()

```

```{r}
#| echo: false
#| fig.cap: "Figure _. The proportion of days open to fishing per month that a roving creel survey was conducted in 2022 and 2023 on the Humptulips River. Surveys were conducted from September 1, 2022 to November 29, 2023. The fishery was closed through emergency regulations from October 8-28, 2022 due to drought conditions. The upper section of the river from the Highway 101 Bridge upstream to the confluence of the East Fork and West Fork Humptulips Rivers was closed from December 1-31, 2022, and the lower section of the river from the mouth to the Highway 101 bridge  was closed from December 17-31, 2022. In December 2023 the fishery was closed by emergency rule."

report_tabs$combined_dates_monthly_p |>
ggplot(aes(x = month, y = proportion_surveyed)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "black") +
  facet_wrap(~ year, ncol = 1, scales = "free_y") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
       x = "Month",
       y = "Proportion of open fishing days surveyed") +
  theme_bw()

```

```{r interview fishing time / angler type assigments}
#| echo: false

# use prep_dwg_interview_fishing_time, prep_dwg_interview_angler_types functions to calculate fishing time and create angler type assignments 
dwg$interview <- 
    CreelEstimateR::prep_dwg_interview_fishing_time(
      dwg_interview = dwg$interview,
      person_count_type = params$person_count_type,
      min_fishing_time = params$min_fishing_time,
      study_design = params$study_design
    )

dwg$interview <- 
    CreelEstimateR::prep_dwg_interview_angler_types(
      interview_fishing_time = dwg$interview,
      study_design = params$study_design,
      boat_type_collapse = params$boat_type_collapse,
      fish_location_determines_type = params$fish_location_determines_type,
      angler_type_kayak_pontoon = params$angler_type_kayak_pontoon
    )
      
```

### Angler interviews

```{r report_interview_totals}
#| echo: false

# summarize the number of interviews obtained by year and angler type 
report_tabs$interview_totals <- dwg$interview |> group_by(year) |> mutate(annual_total_interviews = n_distinct(interview_id)) |> 
  group_by(year, angler_final, annual_total_interviews) |> 
  summarise(
    total_angler_count = sum(angler_count),
    total_individual_count = sum(total_group_count),
    annual_interview_group_count = n_distinct(interview_id)
    )

# summarize the number of interviews obtained by trip_status (complete/incomplete trips)
report_tabs$trip_status_summary <- dwg$interview |> 
  count(year, angler_final, trip_status) |> 
  arrange(year, angler_final, trip_status) |> 
  group_by(year, angler_final) |> 
  mutate(
    total_interviews = sum(n),
    trip_status_proportion = round(n / total_interviews, 3) * 100
  )

# pull out the total number of interviews obatined in section 2
report_tabs$section_2_interviews <- dwg$interview |> filter(section_num == 2) |>  count(year, section_num) |> summarise(sum(n)) |> pluck() 

```

In `r report_tabs$interview_totals[1,1]`, a total of `r report_tabs$interview_totals[1,3]` interviews were obtained, representing `r report_tabs$interview_totals[1,6]` `r report_tabs$interview_totals[1,2]` angler groups and `r report_tabs$interview_totals[2,6]` `r report_tabs$interview_totals[2,2]` angler groups. The majority of bank anglers contacted had completed their fishing trips (`r report_tabs$trip_status_summary[1,4]`/`r report_tabs$trip_status_summary[1,5]`; `r report_tabs$trip_status_summary[1,6]`%). Virtually all boat anglers contacted had also completed their angling trip (`r report_tabs$trip_status_summary[3,4]`/`r report_tabs$trip_status_summary[3,5]`; `r report_tabs$trip_status_summary[3,6]`%)

In `r report_tabs$interview_totals[3,1]`, a total of `r report_tabs$interview_totals[3,3]` interviews were obtained, representing `r report_tabs$interview_totals[3,6]` `r report_tabs$interview_totals[3,2]` angler group interviews and `r report_tabs$interview_totals[4,6]` `r report_tabs$interview_totals[4,2]` angler group interviews. The vast majority of interviews were obtained in section 1. Again, the majority of bank anglers contacted had completed their fishing trips (`r report_tabs$trip_status_summary[5,4]`/`r report_tabs$trip_status_summary[5,5]`; `r report_tabs$trip_status_summary[5,6]`%), but sampling in 2023 reflected an increase the number of bank anglers contacted mid-trip. Similar to sampling in 2022, virtually all boat anglers contacted had also completed their angling trip (`r report_tabs$trip_status_summary[7,4]`/`r report_tabs$trip_status_summary[7,5]`; `r report_tabs$trip_status_summary[7,6]`%). Across 2022 and 2023 very few interviews were obtained in section 2 (*n* = `r report_tabs$section_2_interviews`), representing the upper river from the Highway 101 bridge upstream to the confluence of the East Fork and West Fork Humptulips Rivers, due to limited sampler access (Table \_).

```{r interview_and_catch}
#| echo: false

dwg$interview |>
  count(water_body, year, section_num, angler_final) |> 
  arrange(year, section_num, angler_final) |> 
  select(year, section_number = section_num, angler_type = angler_final, count = n) |> 
  rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence()) |> 
  gt(caption = "Table _. The number of interviews obtained during roving creel surveys, grouped by year, section, and angler type.") 

```

```{r}
#| echo: false
#| include: false

# summary stats describing interview sampling effort by year and month, with daily minimum, maximum, and mean number of interviews obtained
dwg$interview |>
  group_by(year, month, event_date) |>
  summarise(
    daily_count = n()
  ) |>
  group_by(year, month) |>
  summarise(
    min = min(daily_count),
    max = max(daily_count),
    mean = mean(daily_count),
    sd = sd(daily_count)
    ,
    .groups = "drop"
  ) |> 
  gt(caption = "Summary stats describing angler sampling by year and month, with daily minimum, maximum, and mean number of interviews obtained.") 

```

```{r}
#| echo: false
#| include: false

# number of individuals within groups interviewed
 dwg$interview |>
  group_by(year, month, angler_final) |>
  summarise(
    total_angler_count = sum(angler_count),
    total_individual_count = sum(total_group_count)
    ) |>
  rename(angler_type = angler_final) |>
  rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence()) |>
  ggplot(aes(x = as.factor(Month), y = `Total individual count`, fill = `Angler type`)) +
  geom_bar(stat = "identity", position = position_stack(), color = "black") +
  facet_wrap(~ Year, scales = "free_x") +
  scale_x_discrete(limits = c('9', '10', '11', '12')) +
  scale_fill_manual(values = c("bank" = "steelblue", "boat" = "orange"),
                    labels = c("Bank", "Boat")) +
  labs(
       x = "Month",
       y = "Number of angler group members interviewed",
       fill = "Angler type") +
  theme_bw()

```

```{r}
#| echo: FALSE
#| fig.cap: "Figure _. The number of interviews obtained, categorized by bank (blue) and boat (orange) angling groups, per month on the Humptulips River from September 1, 2022 to November 29, 2023. The fishery was closed through emergency regulations from October 8-28, 2022 due to drought conditions. The upper section of the river from the Highway 101 Bridge upstream to the confluence of the East Fork and West Fork Humptulips Rivers was closed from December 1-31, 2022, and the lower section of the river from the mouth to the Highway 101 bridge  was closed from December 17-31, 2022. In December 2023 the fishery was closed by emergency rule."


# number of angler groups interviewed (equivalent to # of interviews obtained)
 dwg$interview |> 
  count(year, month, angler_final, name = "interview_count") |> # count of interviews 
  rename(angler_type = angler_final) |>
  rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence()) |> 
  ggplot(aes(x = as.factor(Month), y = `Interview count`, fill = `Angler type`)) +
  geom_bar(stat = "identity", position = position_stack(), color = "black") +
  facet_wrap(~ Year, scales = "free_x") +
  scale_x_discrete(limits = c('9', '10', '11', '12')) +
  scale_fill_manual(values = c("bank" = "steelblue", "boat" = "orange"), 
                    labels = c("Bank", "Boat")) +
  labs(
       x = "Month",
       y = "Number of angler group interviews",
       fill = "Angler type") +
  theme_bw()

```

```{r, month_mean_trip_length}
#| echo: FALSE
 
report_tabs$mean_trip_length_monthly <- dwg$interview |> 
  filter(trip_status == "Complete", previously_interviewed == 0) |>
  group_by(year, month, angler_final) |> 
  summarize(
    n_completed_angler_trips = n(),
    mean_trip_length = mean(fishing_time),
    sd = sd(fishing_time),
    se = sd / sqrt(n_completed_angler_trips),
    cv = sd / mean_trip_length
  ) |> 
  mutate(
    mutate(across(mean_trip_length:cv, ~ round(., 2)))
  ) |> 
  rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence()) |>
  rename(`Angler type` = `Angler final`, SD = Sd, SE = Se, CV = Cv)

```

```{r}
#| echo: FALSE
#| fig.cap: "Figure _. Mean monthly trip time (± SD) of bank (blue) and boat (orange) angling groups surveyed on the Humptulips River during 2022 and 2023. Surveys were conducted from September 1, 2022 to November 29, 2023. The fishery was closed through emergency regulations from October 8-28, 2022 due to drought conditions. The upper section of the river from the Highway 101 Bridge upstream to the confluence of the East Fork and West Fork Humptulips Rivers was closed from December 1-31, 2022, and the lower section of the river from the mouth to the Highway 101 bridge  was closed from December 17-31, 2022. In December 2023 the fishery was closed by emergency rule."

# Generate bar plot with error bars
report_tabs$mean_trip_length_monthly |> 
ggplot(aes(x = factor(Month), y = `Mean trip length`, fill = `Angler type`)) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  geom_errorbar(aes(ymin = `Mean trip length` - SD, ymax = `Mean trip length` + SD), 
                position = position_dodge(0.9), width = 0.25) +
  labs(
    x = "Month",
    y = "Mean trip length",
    fill = "Angler type"
  ) +
  facet_wrap(~ Year, scales = "free_x") +
  scale_x_discrete(limits = c('9', '10', '11', '12')) +
  scale_fill_manual(values = c("bank" = "steelblue", "boat" = "orange"), 
                    labels = c("Bank", "Boat")) +
  theme_bw()

```

```{r catch}
#| echo: false
#| include: false

# raw reported catch tables. add this information to catch estimates, don't need to take up report space for distinct raw catch tables!

dwg$interview_catch <- dwg$interview |>
  left_join(dwg$catch, by = "interview_id") # one to many as one interview can link to one or more catch records

dwg$interview_catch <- dwg$interview_catch |> 
dplyr::select(-catch_group) |> 
  mutate(
    fin_mark = case_when(
      fin_mark == "AD" ~ "Adipose fin clipped",
      fin_mark == "UM" ~ "Unmarked",
      fin_mark == "UNK" ~ "Unknown",
      TRUE ~ fin_mark
    )) |> 
  unite(
    col = catch_group,
    species, life_stage, fin_mark, fate,
    sep = "_",
    remove = FALSE
  ) 

```

```{r in-line catch table Chinook}
#| echo: false
#| results: asis
#| include: false

report_tabs$Chinook_raw_catches <- dwg$interview_catch |> 
  dplyr::filter(!is.na(fish_count)) |>
  mutate(fin_mark = fct_relevel(fin_mark, "Adipose fin clipped", "Unmarked", "Unknown")) |> 
  dplyr::filter(species == "Chinook", life_stage == "Adult") |>
  dplyr::group_by(year, species, life_stage, fin_mark, fate) |>
  dplyr::summarise(count = sum(fish_count), .groups = "drop") |> 
  arrange(year, species, life_stage, fin_mark, fate) |> 
  rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence())

# Split data by year and generate gt tables
tables_by_year_Chinook <- report_tabs$Chinook_raw_catches |>
  group_split(Year) |>
  map(function(df) {
    df |>
      gt(caption = paste("Total catches of Chinook salmon in", unique(df$Year), "reported during angler interviews. Note that these data represent raw catches and are not expanded for angler effort. See appendix _ for a full table of reported Chinook catches by life stage, fin mark, and fate."))
  })

# map(tables_by_year_Chinook, ~ print(.x)) #this doesn't give same rendered results as manual selections below:

tables_by_year_Chinook[[1]]

tables_by_year_Chinook[[2]]

```

```{r in-line catch table Coho}
#| include: false

report_tabs$coho_raw_catches <- dwg$interview_catch |>
  dplyr::filter(!is.na(fish_count)) |>
  mutate(fin_mark = fct_relevel(fin_mark, "Adipose fin clipped", "Unmarked", "Unknown")) |> 
  dplyr::filter(species == "Coho", life_stage == "Adult") |>
  dplyr::group_by(year, species, life_stage, fin_mark, fate) |>
  dplyr::summarise(count = sum(fish_count), .groups = "drop") |>
  arrange(year, species, life_stage, fin_mark, fate) |>
  rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence())

# Split data by year and generate gt tables
tables_by_year_Coho <- report_tabs$coho_raw_catches |>
  group_split(Year) |>
  lapply(function(df) {
    df |>
      gt(caption = paste("Total catches of Coho salmon in", unique(df$Year), "reported during angler interviews and summarized by life stage, mark status, and fate. Note that these data represent raw catches and are not expanded for angler effort."))
  })

tables_by_year_Coho[[1]]

tables_by_year_Coho[[2]]

```

# Appendix

{{< pagebreak >}}

```{r appendix catch tables}
#| echo: false
#| eval: false
#| include: false

# dwg$interview_catch |> 
#   dplyr::filter(!is.na(fish_count)) |>
#   dplyr::filter(species == "Chinook") |>
#   # dplyr::select(catch_group, fish_count) |>
#   dplyr::group_by(year, species, life_stage, fin_mark, fate) |>
#   dplyr::summarise(count = sum(fish_count), .groups = "drop") |> 
#   arrange(year, species, life_stage, fin_mark, fate) |> 
#   rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence()) |>
#   gt(caption = "Total catches of Chinook salmon reported during angler interviews and summarized by life stage, mark status, and fate. Note that these data represent raw catches and are not expanded for angler effort.")
# 
# # catches of Chinook are way higher in 2021, if we look at coho do we see a similar pattern?
# 
# dwg$interview_catch |> 
#   dplyr::filter(!is.na(fish_count)) |>
#   dplyr::filter(species == "Coho") |>
#   # dplyr::select(catch_group, fish_count) |>
#   dplyr::group_by(year, species, life_stage, fin_mark, fate) |>
#   dplyr::summarise(count = sum(fish_count), .groups = "drop") |> 
#   arrange(year, species, life_stage, fin_mark, fate) |> 
#   rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence()) |>
#   gt(caption = "Total catches of Coho salmon reported during angler interviews and summarized by life stage, mark status, and fate. Note that these data represent raw catches and are not expanded for angler effort.")

```

```{r}
#| echo: false
#| eval: false
#| include: false

data_processed <- dwg$interview_catch |> 
  mutate(fin_mark = fct_relevel(fin_mark, "Adipose fin clipped", "Unmarked", "Unknown")) |> 
  dplyr::filter(!is.na(fish_count)) |>
  dplyr::filter(species == "Coho") |>
  dplyr::group_by(year, species, life_stage, fin_mark, fate) |>
  dplyr::summarise(count = sum(fish_count), .groups = "drop") |> 
  arrange(year, species, life_stage, fin_mark, fate) |> 
  rename_with(~ . |>  str_replace_all("_", " ") |> str_to_sentence())

# Split data by year and generate gt tables
tables_by_year_coho_app <- data_processed |> 
  group_split(Year) |>
  lapply(function(df) {
    df |>
      gt(caption = paste("Total catches of Coho salmon in", unique(df$Year), "reported during angler interviews and summarized by life stage, mark status, and fate. Note that these data represent raw catches and are not expanded for angler effort."))
  })

tables_by_year_coho_app[[1]]
tables_by_year_coho_app[[2]]

```
