---
title: "Study area maps"
author: "Evan Booher"
editor: visual
---

# load packages

```{r}
library("tidyverse")
library("ggmap")
library("gt")
```

# load local spatial data from repo

```{r}
site_coords <- readr::read_csv("../cleaned_data/site_points.csv") |> 
  rename(lon = long)
section_coords <- readr::read_csv("../cleaned_data/section_points.csv")

```

# pull data from data.wa.gov

```{r}
# define fisheries we want data for
fisheries <- c("Nisqually salmon 2021", "Nisqually salmon 2022", "Nisqually salmon 2023")

# download the data from data.wa.gov
datapull <- set_names(fisheries) |> 
  map(~CreelEstimateR::fetch_dwg(fishery_name = .x))

# bind together the fishery_location_lut's for each year
fishery_location_luts <- datapull$`Nisqually salmon 2021`$fishery_manager |> 
  bind_rows(datapull$`Nisqually salmon 2022`$fishery_manager, datapull$`Nisqually salmon 2023`$fishery_manager) |>
  mutate(
    year = str_extract(fishery_name, "\\d{4}$")
  )
  
```

# condensed table of locations used across years

```{r}

#index count locations across years 
fishery_location_luts |> filter(location_type == "Site") |>  select(location_code, year) |> pivot_wider(names_from = year, values_from = year, values_fill = NA) |>
  mutate(
    across(.cols = -location_code, ~ if_else(!is.na(.), TRUE, FALSE))
  ) |> 
  gt()


```

# map of index effort count sites

```{r}
ggmap::register_stadiamaps("7dba3f5b-598a-41b1-bef8-aea60fb3ce1b")

qmplot(lon, lat, data = site_coords, maptype = "stamen_toner_lite", color = I("red"))



```

## next steps

### add sections as spatial lines

### facet for a map for each year

### in plain language describe the choices made for section numbers and how we should refactor for comparability in three year review

### evaluate the fishery_location_lut across three years - what was consistent and what changed across years?

#### tank crossing - only accessible for angler interviews in 2021
