---
title: "Study area maps"
author: "Evan Booher"
editor: visual
---

# load packages

```{r}
library("ggmap")
```

# load local spatial data from repo

```{r}
site_coords <- readr::read_csv("../cleaned_data/site_points.csv")
section_coords <- readr::read_csv("../cleaned_data/section_points.csv")

```

# pull data from data.wa.gov
```{r}
# define fisheries we want data for
fisheries <- c("Nisqually salmon 2021", "Nisqually salmon 2022", "Nisqually salmon 2023")

# download the data from data.wa.gov
datapull <- set_names(fisheries) |> 
  map(~CreelEstimateR::fetch_dwg(fishery_name = .x))

# bind together the fishery_location_lut's for each year
fishery_location_luts <- datapull$`Nisqually salmon 2021`$fishery_manager |> 
  bind_rows(datapull$`Nisqually salmon 2022`$fishery_manager, datapull$`Nisqually salmon 2023`$fishery_manager)

```

# map of index effort count sites
```{r}
ggmap::register_stadiamaps("7dba3f5b-598a-41b1-bef8-aea60fb3ce1b")

# Get the satellite basemap 
basemap <- get_map(location = c(long = mean(c(bbox["xmin"], bbox["xmax"])), lat = mean(c(bbox["ymin"], bbox["ymax"]))), zoom = 12, maptype = "stamen_terrain", source = "stadia")

# Create a simple map using ggplot2
ggmap(basemap) +
  geom_point(data = site_coords, color = "blue") + 
  labs(title = "Nisqually River creel effort index sites")
```

## next steps
### add sections as spatial lines
### facet for a map for each year 
### in plain language describe the choices made for section numbers and how we should refactor for comparability in three year review 
### evaluate the fishery_location_lut across three years - what was consistent and what changed across years?
#### tank crossing - only accessible for angler interviews in 2021

